---
title: "Prelim Analysis/Data Inspection"
author: "Daniel Shephard, Jake Bowers, Michael DiDomenico"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_document:
      graphics: yes
      fig_caption: yes
      fig_height: 4
      fig_width: 4
---



```{r include=FALSE, cache=FALSE}
# Some customization.  You can alter or delete as desired (if you know what you are doing).
# knitr settings to control how R chunks work.

require(knitr)
opts_chunk$set(tidy=FALSE,     # display code as typed
	       size="small",    # slightly smaller font for code
	       echo=FALSE,
	       results='markup',
	       strip.white=TRUE,
	       ## fig.path='figs/fig',
	       cache=FALSE,
	       highlight=TRUE,
	       width.cutoff=132,
	       size='footnotesize',
	       out.width='.9\\textwidth',
	       message=FALSE,
	       comment=NA)

options(width=132,scipen=8,digits=4)
```


# Overview of the Project


## Goal:

Accelerate finding employment and reduce time on unemployment insurance among unemployment insurance clients in Oregon. To test behaviorally informed changes that can be scaled up to other states and incorporated into DOL guidance.

This file executes the analysis of the experiment. It relies on files produced by the `datasetup.Rmd` file.

```{r loadlibs}
library(lmtest)
library(sandwich)
library(car)
library(NPC)
library(RItools)
library(coin)
library(dplyr)
```

The two working data files are produced by the `datasetup.Rmd` file.

```{r}
load("Data/or.df.TOTAL.rda")
load("Data/monthlyordat.rda") ## Contains the month by month data
```


```{r Welcomes_by_Office, echo=FALSE}

#Create a table showing the number of welcome meetings by month for each office
#An important NOTE that welcome meetings are not equivalent to observations
#There can be someone that applied for UI and in the sample but has not done their welcome
#If this is the case then they have not received a payment, b/c the welcome is required
offices <- c(unique(or.df.TOTAL$'277.Field.Office'))[1:14]
months <- c('mar', 'apr', 'may', 'jun', 'jul', 'aug', 'sep')
desc.table.1 <- matrix(nrow = 14, ncol = 7,
		       dimnames = list(offices, months))

#Add the counts of welcomes for March
for(i in 1:14)
{
	desc.table.1[i, 1] <- length(which(or.df1$`277.Field.Office`==offices[i]))
}

#Add the counts of welcomes for April
for(i in 1:14)
{
	desc.table.1[i, 2] <- length(which(or.df2$`277.Field.Office`==offices[i]))
}

#Add the counts of welcomes for May
for(i in 1:14)
{
	desc.table.1[i, 3] <- length(which(or.df3$`277.Field.Office`==offices[i]))
}

#Add the counts of welcomes for June
for(i in 1:14)
{
	desc.table.1[i, 4] <- length(which(or.df4$`277.Field.Office`==offices[i]))
}

#Add the counts of welcomes for July
for(i in 1:14)
{
	desc.table.1[i, 5] <- length(which(or.df5$`277.Field.Office`==offices[i]))
}

#Add the counts of welcomes for August
for(i in 1:14)
{
	desc.table.1[i, 6] <- length(which(or.df6$`277.Field.Office`==offices[i]))
}

#Add the counts of welcomes for September
for(i in 1:14)
{
	desc.table.1[i, 7] <- length(which(or.df7$`277.Field.Office`==offices[i]))
}

```

## Table of welcome conversations by field office

The following table shows the number of welcome conversations conducted at each of the field offices in each month. The first column has the field office labels.

```{r}
knitr::kable(desc.table.1)
```


# The Experiment:

The experiment randomly assigned the new set of procedures to 7 out of the 14
offices after first grouping the offices into pairs.  Pairs were created by
both urban, geography, size (number of people who came for UI), and quality of
staff (judgement of the management in Oregon). Pairs provided by the management
of the Oregon State program.

```{r Pull_Pairs, results='hide'}
#----ORIGINAL CODE FOR RANDOMIZATION------
##pairfilepath <- "C:/Users/DanielDShephard/Documents/SBST/DOL/UI Oregon/Analysis/OR_Analysis/Oregon_Pairs_14-Field-Offices_20160111.csv" 
pairfilepath <- "Data/Oregon_Pairs_14-Field-Offices_20160111.csv" 
pairs2 <- read.csv(pairfilepath, row.names = 1)
row.names(pairs2)<-pairs2$fo.num
pairs2 <- pairs2[order(pairs2$pair), ] #ensure that all pairs are next to each other
#Random generation of uniform distribution x14, one number per field office
#The highest number in the pair is allocated to the treatment group
pairs2$treat <- 0
set.seed(20160108) #note, ran once with 20150108, but changed to match date
ran.no <- runif(14)
for (i in 1:7)
{
	if(ran.no[i*2] > ran.no[(i*2)-1])
	{
		pairs2[i*2, "treat"] <- 1
	}
	else
	{
		pairs2[i*2-1, "treat"] <- 1
	}

}

#Check the frequency of assignment with different seeds to ensure 50% chance
#Run allocation logic used above through 1000 different seeds
pair2 <- NA
pair2 <- as.numeric(pair2)
for (q in 1:1000) #run 1000 simulations of different seeds
{
	set.seed(20160107+q)
	ran.no <- runif(14)
	for (i in 1:7) #for a set of 7 pairs with the current seed do the following
	{
		if(ran.no[i*2] > ran.no[(i*2)-1]) #if the even number in the pair is greater than the odd number
		{
			pair2 <- rbind(pair2, 1) #then give it a treatment value of 1
		}
		else
		{
			pair2 <- rbind(pair2, 0) #else give the second in the pair a value of 0
		}
	}
}
treat.probability <- mean(pair2, na.rm=TRUE)
samp.m1 <- (lm(ui.wks ~ treat, data=pairs2)) #original scores
samp.m2 <- (lm(tot ~ treat, data=pairs2)) #original scores
summary(samp.m1)
summary(samp.m2)
```

```{r pairstable}
#Write out the new dataset with treatment allocation
#The number of welcomes and average weeks on UI is from 2015 data (Pre Intervention)
knitr::kable(pairs2[order(pairs2$pair, pairs2$treat), c(1:2, 14, 9, 12, 10)]) #Order by pairs, with control pair office first

```

### Outcomes

The overall outcomes collected at the individual level within the period
covered by the data pull included:

 - `welcome` Whether the person had a welcome meeting
 - `uiexit` Whether the person stopped making unemployement claims for 3 weeks in a row.
 - `employed` Whether a different company reports wages for the person in a quarter after the quarter they applied for UI
 - `time.to.exit` The number of ?days? between making the first claim and exiting the program.
 - `plans` Number of action plans uploaded by an individual in the program
 - `wages` The wages earned by the person (?average? ?weekly?)

??Why is time.to.exit negative sometimes? (DS: What is the frequency of this? Are they only clustered at the very start of the trial? I could imagine this happening in the March group.)

```{r results=FALSE}
outcomes <- c("welcome","uiexit","employed","time.to.exit","plans","wages")
summary(or.df.TOTAL[,outcomes])
aggvars <- c("insert",outcomes)
```

## Employment Center Level analysis

Because the intervention was randomized at the level of the employment office, we compare outcomes at that level: did offices in the new program display outcomes that differed from the offices operating under the status quo program?

```{r}
## This next creates more columns than I would like
officedat <- or.df.TOTAL %>% group_by(fo.num) %>% summarize_at(aggvars,.funs=funs(mean(.,na.rm=TRUE),sum(is.na(.)),n()))
officedf <- as.data.frame(officedat[,c("fo.num",grep("mean",names(officedat),value=TRUE),"wages_sum","insert_n")])
row.names(officedf)<-officedf$fo.num
names(officedf)[names(officedf)==c("wages_sum","insert_n")] <- c("missingwage","N")

## Can Mike double check tha the means per office are correct and can we report out a table of this? Just want to double check. :)
#Also, is "wages_sum" the number of NAs for the wage outcome?
```


```{r}
## Add treatment assignment info
officedf$fo.num <- as.numeric(officedf$fo.num)
wrkdat <- inner_join(officedf,pairs2,by="fo.num")
wrkdat <- wrkdat[order(wrkdat$pair,wrkdat$treat),]
```


```{r}
## Pair-mean center to handle the pairing:

wrkdat$treatMD <- with(wrkdat,treat - ave(treat,pair))
for(i in outcomes){
	message(i)
	nm <- paste(i,"_mean",sep="")
	wrkdat[,paste(i,"MD",sep="")]<- wrkdat[[nm]] - ave(wrkdat[[nm]],wrkdat$pair)
}
knitr::kable(wrkdat) #If this table is too big it could be broken into pieces
#This looks right to me, but I don't 100% follow the coding. Can Mike double check the pair centering?


```

# Office Level Effects

```{r estimates, results='hide'}
## First, just show two equivalent ways to get the ITT
welcomeATEa <- lm(welcome_mean~treat+pair,data=wrkdat)
welcomeATE <- lm(welcomeMD~treatMD,data=wrkdat)
stopifnot(all.equal(coef(welcomeATEa)[[2]],coef(welcomeATE)[[2]]))

## Next calc the diff of means for all outcomes
aggoutcomes <- paste(outcomes,"MD",sep="")
theITTS <- sapply(aggoutcomes,function(v){
			  coef(lm(reformulate("treatMD",response=v),data=wrkdat))[[2]]
})
theITTS
```


```{r tests, results='hide'}
## The coin package allows us to use permutations/repetitions of the experiment to assess the hypothesis of no effects in a block-randomized experiment
wrkdat$treatF <- factor(wrkdat$treat,levels=c("1","0"))
set.seed(20161102)
welcomeTest <- oneway_test(welcome_mean ~ treatF | pair, data=wrkdat, distribution=approximate(B=5000))
welcomeTest

## The following uses a large sample approximation to the randomization distribution (Mostly just a check on the preceding)
xb1<-xBalance(treat ~ welcome_mean,strata=list(pair=~pair),data=wrkdat,report="all")
xb1$results

## Now do it for all
theps <- sapply(aggoutcomes,function(v){
			 results <- oneway_test(reformulate("treatF | pair ",response=v),data=wrkdat,distribution=approximate(B=5000))
			 return(pvalue(results))
})
theps
```


We present differences of means below --- they are an unbiased esimtator of the average treatment effect. To assess the amount of information we have against the null of no effects, we present the results of hypothesis tests.Because the sample size is so small, we use a randomization-based approach.

```{r}
cbind(theITTS,theps)
```

# Test the null of no effects for any outcome

Here we see that we can argue against the null of no effects for any outcome with a p-value of roughly .09. I also show the individual p-values here because they are exact (as compared to the approximative ones above). Here using the Student's T test statistic.

```{r}
npcresults <- NPC(wrkdat,tr.var="treat",tr.label="1",y.vars=paste(outcomes,"_mean",sep=""),block.var="pair",alternative="two.sided",return.matrix=TRUE,FWE.adj=FALSE,print.steps=FALSE)
npcresults$p.values
```


