---
title: "Prelim Analysis/Data Inspection"
author: "Daniel Shephard, Jake Bowers, Michael DiDomenico"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_document:
      graphics: yes
      fig_caption: yes
---



```{r include=FALSE, cache=FALSE}
# Some customization.  You can alter or delete as desired (if you know what you are doing).
# knitr settings to control how R chunks work.

require(knitr)
opts_chunk$set(tidy=FALSE,     # display code as typed
	       size="small",    # slightly smaller font for code
	       echo=FALSE,
	       results='markup',
	       strip.white=TRUE,
	       ## fig.path='figs/fig',
	       cache=FALSE,
	       highlight=TRUE,
	       width.cutoff=132,
	       size='footnotesize',
	       #out.width='.9\\textwidth',
	       message=FALSE,
	       comment=NA)

options(width=132,scipen=8,digits=4)
```


# Overview of the Project


## Goal:

Accelerate finding employment and reduce time on unemployment insurance among unemployment insurance clients in Oregon. 

This file executes the analysis of the experiment. It relies on files produced by the `datasetup.Rmd` file.

```{r loadlibs}
library(lmtest)
library(sandwich)
library(car)
library(NPC)
library(RItools)
library(coin)
library(dplyr)
library(lubridate)
```

The two working data files are produced by the `datasetup.Rmd` file.

```{r}
load("Data/clean/or.df.rda")

load("Data/clean/monthlyordat.rda") ## Contains the month by month data
```

## Table of welcome conversations by field office

The following table shows the number of welcome conversations conducted at each
of the field offices in each month. Some of these welcomes come from the month
by month data pulls. Others come from the updated wages file. The first column
has the field office labels.

```{r}
tmp <- sapply(split(or.df,months(or.df$claim.date)),function(dat){ table(dat$fonum) })
tmp <- tmp[,c("March","April","May","June","July","August","September","October","November","December")]
```


```{r}
knitr::kable(tmp)

```


# The Experiment:

The experiment randomly assigned the new set of procedures to 7 out of the 14
offices after first grouping the offices into pairs.  Pairs were created by
both urban, geography, size (number of people who came for UI), and quality of
staff (judgement of the management in Oregon). Pairs provided by the management
of the Oregon State program.

```{r Pull_Pairs, results='hide'}
#----ORIGINAL CODE FOR RANDOMIZATION------
##pairfilepath <- "C:/Users/DanielDShephard/Documents/SBST/DOL/UI Oregon/Analysis/OR_Analysis/Oregon_Pairs_14-Field-Offices_20160111.csv"
pairfilepath <- "Data/Oregon_Pairs_14-Field-Offices_20160111.csv"
pairs2 <- read.csv(pairfilepath, row.names = 1)
row.names(pairs2)<-pairs2$fo.num
pairs2 <- pairs2[order(pairs2$pair), ] #ensure that all pairs are next to each other
#Random generation of uniform distribution x14, one number per field office
#The highest number in the pair is allocated to the treatment group
pairs2$treat <- 0
set.seed(20160108) #note, ran once with 20150108, but changed to match date
ran.no <- runif(14)
for (i in 1:7)
{
	if(ran.no[i*2] > ran.no[(i*2)-1])
	{
		pairs2[i*2, "treat"] <- 1
	}
	else
	{
		pairs2[i*2-1, "treat"] <- 1
	}

}

#Check the frequency of assignment with different seeds to ensure 50% chance
#Run allocation logic used above through 1000 different seeds
pair2 <- NA
pair2 <- as.numeric(pair2)
for (q in 1:1000) #run 1000 simulations of different seeds
{
	set.seed(20160107+q)
	ran.no <- runif(14)
	for (i in 1:7) #for a set of 7 pairs with the current seed do the following
	{
		if(ran.no[i*2] > ran.no[(i*2)-1]) #if the even number in the pair is greater than the odd number
		{
			pair2 <- rbind(pair2, 1) #then give it a treatment value of 1
		}
		else
		{
			pair2 <- rbind(pair2, 0) #else give the second in the pair a value of 0
		}
	}
}
treat.probability <- mean(pair2, na.rm=TRUE)
samp.m1 <- (lm(ui.wks ~ treat, data=pairs2)) #original scores
samp.m2 <- (lm(tot ~ treat, data=pairs2)) #original scores
summary(samp.m1)
summary(samp.m2)
```

The number of welcomes and average weeks on UI is from 2015 data (Pre Intervention)

```{r pairstable}
#Write out the new dataset with treatment allocation
knitr::kable(pairs2[order(pairs2$pair, pairs2$treat), c(1:2, 14, 9, 12, 10)]) #Order by pairs, with control pair office first
```

### Outcomes

The overall outcomes collected at the individual level within the period
covered by the data pull included:

 - `welcome` Whether the person had a welcome meeting
 - `uiexit` Whether the person stopped making unemployment claims for 3 weeks in a row.
 - `employed` Whether a different company reports wages for the person in a quarter after the quarter they applied for UI
 - `time.to.exit` The number of days between making the first claim and exiting the program.
 - `plans` Number of action plans uploaded by an individual in the program
 - `wages` The wages earned by the person (?average? ?weekly?)

??Why is time.to.exit negative sometimes? (DS: What is the frequency of this? Are they only clustered at the very start of the trial? I could imagine this happening in the March group). And yes, the 'time.to.exit' should represent the number of days between the two dates. (PT: Negative times to exit occur in all months, and are always within 30 days, so perhaps an error in data entry?)

```{r results=FALSE}
outcomes <- c("welcome","uiexit","employed","time.to.exit","plans","wages")
summary(or.df[,outcomes])
aggvars <- c("insert",outcomes)
```

```{r}
table(or.df$uiexit,months(or.df$claim.date))
```

## Employment Center Level Analysis

Because the intervention was randomized at the level of the employment office, we compare outcomes at that level: did offices in the new program display outcomes that differed from the offices operating under the status quo program?

```{r}

## This next creates more columns than I would like
officedat <- or.df %>% group_by(fonum) %>% summarize_at(aggvars,.funs=funs(mean(.,na.rm=TRUE),sum(is.na(.)),n()))
officedf <- as.data.frame(officedat[,c("fonum",grep("mean",names(officedat),value=TRUE),"wages_sum","insert_n")])
row.names(officedf)<-officedf$fonum
names(officedf)[names(officedf)==c("wages_sum","insert_n")] <- c("missingwage","N")

## wages_sum is the number of people with missing wages data

## Check that the calculation of the mean within office is correct by looking at one office
stopifnot(all.equal(with(or.df,mean(welcome[fonum=="210"])),officedf["210","welcome_mean"]))
```


```{r mergeontreatment}
## Add treatment assignment info
officedf$fonum <- as.numeric(officedf$fonum)
pairs2$fonum <- pairs2$fo.num
wrkdat <- inner_join(officedf,pairs2,by="fonum")
wrkdat <- wrkdat[order(wrkdat$pair,wrkdat$treat),]
row.names(wrkdat) <- as.character(wrkdat$fonum)
```


```{r}
## Pair-mean center to handle the pairing:

wrkdat$treatMD <- with(wrkdat,treat - ave(treat,pair))
for(i in outcomes){
	message(i)
	nm <- paste(i,"_mean",sep="")
	wrkdat[,paste(i,"MD",sep="")]<- wrkdat[[nm]] - ave(wrkdat[[nm]],wrkdat$pair)
}
knitr::kable(wrkdat) #If this table is too big it could be broken into pieces

## Check the pair centering
stopifnot(all.equal(wrkdat$welcomeMD[wrkdat$pair=="a"],
		    with(wrkdat[wrkdat$pair=="a",], welcome_mean - mean(welcome_mean))
		    ) )
```

# Office Level Effects

```{r estimates, results='hide'}
## First, just show two equivalent ways to get the ITT
welcomeATEa <- lm(welcome_mean~treat+pair,data=wrkdat)
welcomeATE <- lm(welcomeMD~treatMD,data=wrkdat)
stopifnot(all.equal(coef(welcomeATEa)[[2]],coef(welcomeATE)[[2]]))

## Next calc the diff of means for all outcomes
aggoutcomes <- paste(outcomes,"MD",sep="")
theITTS <- sapply(aggoutcomes,function(v){
			  coef(lm(reformulate("treatMD",response=v),data=wrkdat))[[2]]
})
theITTS
```


```{r tests, results='hide'}
## The coin package allows us to use permutations/repetitions of the experiment to assess the hypothesis of no effects in a block-randomized experiment
wrkdat$treatF <- factor(wrkdat$treat,levels=c("1","0"))
set.seed(20161102)
welcomeTest <- oneway_test(welcome_mean ~ treatF | pair, data=wrkdat, distribution=approximate(B=5000))
welcomeTest

## The following uses a large sample approximation to the randomization distribution (Mostly just a check on the preceding)
xb1<-xBalance(treat ~ welcome_mean,strata=list(pair=~pair),data=wrkdat,report="all")
xb1$results

## Now do it for all
theps <- sapply(aggoutcomes,function(v){
			 results <- oneway_test(reformulate("treatF | pair ",response=v),data=wrkdat,distribution=approximate(B=5000))
			 return(pvalue(results))
})
theps
```


We present differences of means below --- they are an unbiased estimator of the average treatment effect. To assess the amount of information we have against the null of no effects, we present the results of hypothesis tests. Because the sample size is so small, we use a randomization-based approach. In short, the intervention appeared effective at increasing the number of plans completed, but did not appear influence any of the remaining outcomes.

```{r}
cbind(theITTS,theps)
```

# Test the null of no effects for any outcome


```{r}
npcresults <- NPC(wrkdat,tr.var="treat",tr.label="1",y.vars=paste(outcomes,"_mean",sep=""),block.var="pair",alternative="two.sided",return.matrix=TRUE,FWE.adj=FALSE,print.steps=FALSE)
npcresults$p.values
```

Here we see that the evidence against the null of no effects for any outcome has a p-value of roughly `r npcresults$p.values[["NPC"]]`. I also show the individual p-values here because they are exact (as compared to the approximate ones above). Here using the Student's T test statistic.

# Effects Excluding Those With a Short Time on the Program

```{r,negativeexit,echo=F,results="hide"}
sum(or.df$time.to.exit<=0)/dim(or.df)[1]
sum(sort(prop.table(with(or.df[or.df$time.to.exit<=0,],table(time.to.exit))),decreasing = T)[1:5])

officedat_noneg <- or.df[or.df$time.to.exit>=0,] %>% group_by(fonum) %>% summarize_at(aggvars,.funs=funs(mean(.,na.rm=TRUE),sum(is.na(.)),n()))
officedf_noneg <- as.data.frame(officedat_noneg[,c("fonum",grep("mean",names(officedat_noneg),value=TRUE),"wages_sum","insert_n")])
row.names(officedf_noneg)<-officedf_noneg$fonum
names(officedf_noneg)[names(officedf_noneg)==c("wages_sum","insert_n")] <- c("missingwage","N")

## Add treatment assignment info
officedf_noneg$fonum <- as.numeric(officedf_noneg$fonum)
pairs2$fonum <- pairs2$fo.num
wrkdat_noneg <- inner_join(officedf_noneg,pairs2,by="fonum")
wrkdat_noneg <- wrkdat_noneg[order(wrkdat_noneg$pair,wrkdat_noneg$treat),]
row.names(wrkdat_noneg) <- as.character(wrkdat_noneg$fonum)

## Pair-mean center to handle the pairing:

wrkdat_noneg$treatMD <- with(wrkdat_noneg,treat - ave(treat,pair))
for(i in outcomes){
	message(i)
	nm <- paste(i,"_mean",sep="")
	wrkdat_noneg[,paste(i,"MD",sep="")]<- wrkdat_noneg[[nm]] - ave(wrkdat_noneg[[nm]],wrkdat_noneg$pair)
}

wrkdat_noneg$treatMD <- with(wrkdat_noneg,treat - ave(treat,pair))
for(i in outcomes){
	message(i)
	nm <- paste(i,"_mean",sep="")
	wrkdat_noneg[,paste(i,"MD",sep="")]<- wrkdat_noneg[[nm]] - ave(wrkdat_noneg[[nm]],wrkdat_noneg$pair)
}

aggoutcomes <- paste(outcomes,"MD",sep="")
theITTS_noneg <- sapply(aggoutcomes,function(v){
			  coef(lm(reformulate("treatMD",response=v),data=wrkdat_noneg))[[2]]
})
wrkdat_noneg$treatF <- factor(wrkdat_noneg$treat,levels=c("1","0"))

theps_noneg <- sapply(aggoutcomes,function(v){
			 results <- oneway_test(reformulate("treatF | pair ",response=v),data=wrkdat_noneg,distribution=approximate(B=5000))
			 return(pvalue(results))
})
npcresults <- NPC(wrkdat_noneg,tr.var="treat",tr.label="1",y.vars=paste(outcomes,"_mean",sep=""),block.var="pair",alternative="two.sided",return.matrix=TRUE,FWE.adj=FALSE,print.steps=FALSE)

```


A small portion of the cases appear to have "negative" times to exiting the program, (N=`r sum(or.df$time.to.exit<=0)` or `r round(sum(or.df$time.to.exit<=0)/dim(or.df)[1]*100,1)` percent of cases)

Time to exiting the program is calculated by the subtracting participant's initial unemployment claim date from the date they exited the unemployment insurance. An individual is considered to have exited the program after three weeks of not claiming benefits. Thus exit dates are always on Sundays when a new claim period begins, while initial claim date may be any day during the week. In fact nearly `r round(sum(sort(prop.table(with(or.df[or.df$time.to.exit<=0,],table(time.to.exit))),decreasing = T)[1:5])*100,1)` percent of these cases have values implying exit dates within seven days of the claim process, suggesting most of the negative cases are due to disconnect between how claims are processed and exit times are calculated. Still, as a robustness check, we repeat the main analysis, removing participants with negative times to exit. The results are essentially unchanged.


```{r}
cbind(theITTS_noneg,theps_noneg)
```



# Effects on Those With More Time in the Program?

One idea is that effects of the program might be smaller among those who have had little time in the program. So, we might inspect the effects among those who have been in the program longer (say, those with `BYE<=201718` (Benefits Year End Date of the 18th week of 2017).

First, however, before we divide up the sample in this way, we should see if there are differential effects of treatment on BYE. Even if we find that the treated and control groups are balanced on this,  if we think that the treatment may have a causal effect on this factor, then, below, we may want to do a different kind of analysis (i.e. consider BYE as the dose and random assignment as the instrument).

```{r byextreat, results="markup"}
## We are just looking for systematic relationships between BYE and treatment status, so it is ok to make this numeric.
or.df$BYEN <- as.numeric(or.df$BYE)

tmp <-  or.df %>% group_by(fonum) %>% dplyr::summarise(bye_mean=mean(BYEN))
wrkdat[as.character(tmp$fonum),"bye_mean"]<-tmp$bye_mean
byeTest <- oneway_test(bye_mean~treatF | pair,data=wrkdat)
byeTest
with(wrkdat,c(mean(bye_mean[treat==1]),mean(bye_mean[treat==0])))
```

The preceding analysis does not suggest different BYE in the treatment and control groups. We now re-do the preceding analysis, but only for the subset of respondents who have spent the most time in the program.


```{r agg2}
## This next creates more columns than I would like
officedat2 <- or.df[or.df$BYE<=201718,] %>% group_by(fonum) %>% summarize_at(aggvars,.funs=funs(mean(.,na.rm=TRUE),sum(is.na(.)),n()))
officedf2 <- as.data.frame(officedat2[,c("fonum",grep("mean",names(officedat2),value=TRUE),"wages_sum","insert_n")])
row.names(officedf2)<-officedf2$fonum
names(officedf2)[names(officedf2)==c("wages_sum","insert_n")] <- c("missingwage","N")

## Add treatment assignment info
officedf2$fonum <- as.numeric(officedf2$fonum)
pairs2$fonum <- pairs2$fo.num
wrkdat2 <- inner_join(officedf2,pairs2,by="fonum")
wrkdat2 <- wrkdat2[order(wrkdat2$pair,wrkdat2$treat),]
row.names(wrkdat2) <- as.character(wrkdat2$fonum)
```


```{r}
## Pair-mean center to handle the pairing:

wrkdat2$treatMD <- with(wrkdat2,treat - ave(treat,pair))
for(i in outcomes){
	message(i)
	nm <- paste(i,"_mean",sep="")
	wrkdat2[,paste(i,"MD",sep="")]<- wrkdat2[[nm]] - ave(wrkdat2[[nm]],wrkdat2$pair)
}
knitr::kable(wrkdat2) #If this table is too big it could be broken into pieces

## Check the pair centering
stopifnot(all.equal(wrkdat2$welcomeMD[wrkdat2$pair=="a"],
		    with(wrkdat2[wrkdat2$pair=="a",], welcome_mean - mean(welcome_mean))
		    ) )
```



```{r estimates2, results='hide'}
theITTS2 <- sapply(aggoutcomes,function(v){
			  coef(lm(reformulate("treatMD",response=v),data=wrkdat2))[[2]]
})
theITTS2
```


```{r tests2, results='hide'}
## The coin package allows us to use permutations/repetitions of the experiment to assess the hypothesis of no effects in a block-randomized experiment
wrkdat2$treatF <- factor(wrkdat2$treat,levels=c("1","0"))
set.seed(20161102)
## Now do it for all
theps2 <- sapply(aggoutcomes,function(v){
			 results <- oneway_test(reformulate("treatF | pair ",response=v),data=wrkdat2,distribution=approximate(B=5000))
			 return(pvalue(results))
})
theps2
```

Here are the estimated ITTs and p-values for the null of no effects among the group that had the most time in the program.

```{r}
cbind(theITTS2,theps2)
```

# Test the null of no effects for any outcome

```{r}
npcresults2 <- NPC(wrkdat2,tr.var="treat",tr.label="1",y.vars=paste(outcomes,"_mean",sep=""),block.var="pair",alternative="two.sided",return.matrix=TRUE,FWE.adj=FALSE,print.steps=FALSE)
npcresults2$p.values[["NPC"]]
```

Here we see that the evidence against the null of no effects for any outcome with a p-value of roughly `r npcresults2$p.values[["NPC"]]`. I also show the individual p-values here because they are exact (as compared to the approximate ones above). The test statistic used here, by default, is Student's T.

## Separate Effects by Month

Below I show the estimated effects by month for each of the six outcomes.

```{r monthly}
# Wrapper Function to calculate by month

monthly_itts<-function(the_month=3){
    ## This next creates more columns than I would like
officedat_month <- or.df[month(or.df$claim.date)==the_month,] %>% group_by(fonum) %>% summarize_at(aggvars,.funs=funs(mean(.,na.rm=TRUE),sum(is.na(.)),n()))
officedf_month <- as.data.frame(officedat_month[,c("fonum",grep("mean",names(officedat_month),value=TRUE),"wages_sum","insert_n")])
row.names(officedf_month)<-officedf_month$fonum
names(officedf_month)[names(officedf_month)==c("wages_sum","insert_n")] <- c("missingwage","N")

## Add treatment assignment info
officedf_month$fonum <- as.numeric(officedf_month$fonum)
pairs2$fonum <- pairs2$fo.num
wrkdat_month <- inner_join(officedf_month,pairs2,by="fonum")
wrkdat_month <- wrkdat_month[order(wrkdat_month$pair,wrkdat_month$treat),]
row.names(wrkdat_month) <- as.character(wrkdat_month$fonum)

## Pair-mean center to handle the pairing:

wrkdat_month$treatMD <- with(wrkdat_month,treat - ave(treat,pair))
for(i in outcomes){
	message(i)
	nm <- paste(i,"_mean",sep="")
	wrkdat_month[,paste(i,"MD",sep="")]<- wrkdat_month[[nm]] - ave(wrkdat_month[[nm]],wrkdat_month$pair)
}

wrkdat_month$treatMD <- with(wrkdat_month,treat - ave(treat,pair))
for(i in outcomes){
	message(i)
	nm <- paste(i,"_mean",sep="")
	wrkdat_month[,paste(i,"MD",sep="")]<- wrkdat_month[[nm]] - ave(wrkdat_month[[nm]],wrkdat_month$pair)
}

if(the_month>9){outcomes<-outcomes[-6]} # no wage data after september
aggoutcomes <- paste(outcomes,"MD",sep="")
theITTS_month <- sapply(aggoutcomes,function(v){
			  coef(lm(reformulate("treatMD",response=v),data=wrkdat_month))[[2]]
})
wrkdat_month$treatF <- factor(wrkdat_month$treat,levels=c("1","0"))

theps_month <- sapply(aggoutcomes,function(v){
			 results <- oneway_test(reformulate("treatF | pair ",response=v),data=wrkdat_month,distribution=approximate(B=5000))
			 return(pvalue(results))
})

return(data.frame(theITTS_month,theps_month,the_month,outcomes))
}

march_itt<-monthly_itts(the_month=3)
april_itt<-monthly_itts(the_month=4)
may_itt<-monthly_itts(the_month=5)
june_itt<-monthly_itts(the_month=6)
july_itt<-monthly_itts(the_month=7)
august_itt<-monthly_itts(the_month=8)
september_itt<-monthly_itts(the_month=9)
october_itt<-monthly_itts(the_month=10)
november_itt<-monthly_itts(the_month=11)
december_itt<-monthly_itts(the_month=12)

monthly_df<-rbind(march_itt,april_itt,may_itt,june_itt,july_itt,august_itt,september_itt,october_itt,november_itt,december_itt)
monthly_df$outcomes<-factor(monthly_df$outcomes)
library(ggplot2)       
montly_plot<-ggplot(monthly_df,aes(x=the_month,y=theITTS_month))+geom_line()+facet_wrap(~outcomes,scales = "free")+labs(x="Month",y="ITT")


```

```{r,results="markup"}
montly_plot
```


