---
title: "Prelim Analysis/Data Inspection"
author: "Daniel Shephard, Jake Bowers, Michael DiDomenico"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_document:
      graphics: yes
      fig_caption: yes
---



```{r include=FALSE, cache=FALSE}
# Some customization.  You can alter or delete as desired (if you know what you are doing).
# knitr settings to control how R chunks work.

require(knitr)
opts_chunk$set(tidy=FALSE,     # display code as typed
	       size="small",    # slightly smaller font for code
	       echo=FALSE,
	       results='markup',
	       strip.white=TRUE,
	       ## fig.path='figs/fig',
	       cache=FALSE,
	       highlight=TRUE,
	       width.cutoff=132,
	       size='footnotesize',
	       #out.width='.9\\textwidth',
	       message=FALSE,
	       comment=NA)

options(width=132,scipen=8,digits=4)
```


# Overview of the Project


## Goal:

Accelerate finding employment and reduce time on unemployment insurance among unemployment insurance clients in Oregon. 

This file executes the analysis of the experiment. It relies on files produced by the `datasetup.Rmd` file.

```{r loadlibs}
library(lmtest)
library(sandwich)
library(car)
library(NPC)
library(RItools)
library(coin)
library(dplyr)
library(lubridate)
library(mosaic)
```

The two working data files are produced by the `datasetup.Rmd` file.

```{r}
load("Data/clean/or.df.rda")
load("Data/clean/monthlyordat.rda") ## Contains the month by month data
```

## Table of welcome conversations by field office

The following table shows the number of welcome conversations conducted at each
of the field offices in each month. Some of these welcomes come from the month
by month data pulls. Others come from the updated wages file. The first column
has the field office labels.

```{r}
tmp <- sapply(split(or.df,months(or.df$claim.date)),function(dat){ table(dat$fonum) })
tmp <- tmp[,c("March","April","May","June","July","August","September","October","November","December")]
```


```{r}
knitr::kable(tmp)

```


# The Experiment:

The experiment randomly assigned the new set of procedures to 7 out of the 14
offices after first grouping the offices into pairs.  Pairs were created by
both urban, geography, size (number of people who came for UI), and quality of
staff (judgement of the management in Oregon). Pairs provided by the management
of the Oregon State program.

```{r Pull_Pairs, results='hide'}
#----ORIGINAL CODE FOR RANDOMIZATION------
##pairfilepath <- "C:/Users/DanielDShephard/Documents/SBST/DOL/UI Oregon/Analysis/OR_Analysis/Oregon_Pairs_14-Field-Offices_20160111.csv"
pairfilepath <- "Data/Oregon_Pairs_14-Field-Offices_20160111.csv"
pairs2 <- read.csv(pairfilepath, row.names = 1)
row.names(pairs2)<-pairs2$fo.num
pairs2 <- pairs2[order(pairs2$pair), ] #ensure that all pairs are next to each other
#Random generation of uniform distribution x14, one number per field office
#The highest number in the pair is allocated to the treatment group
pairs2$treat <- 0
set.seed(20160108) #note, ran once with 20150108, but changed to match date
ran.no <- runif(14)
for (i in 1:7)
{
	if(ran.no[i*2] > ran.no[(i*2)-1])
	{
		pairs2[i*2, "treat"] <- 1
	}
	else
	{
		pairs2[i*2-1, "treat"] <- 1
	}

}

#Check the frequency of assignment with different seeds to ensure 50% chance
#Run allocation logic used above through 1000 different seeds
pair2 <- NA
pair2 <- as.numeric(pair2)
for (q in 1:1000) #run 1000 simulations of different seeds
{
	set.seed(20160107+q)
	ran.no <- runif(14)
	for (i in 1:7) #for a set of 7 pairs with the current seed do the following
	{
		if(ran.no[i*2] > ran.no[(i*2)-1]) #if the even number in the pair is greater than the odd number
		{
			pair2 <- rbind(pair2, 1) #then give it a treatment value of 1
		}
		else
		{
			pair2 <- rbind(pair2, 0) #else give the second in the pair a value of 0
		}
	}
}
treat.probability <- mean(pair2, na.rm=TRUE)
samp.m1 <- (lm(ui.wks ~ treat, data=pairs2)) #original scores
samp.m2 <- (lm(tot ~ treat, data=pairs2)) #original scores
summary(samp.m1)
summary(samp.m2)
```

The number of welcomes and average weeks on UI is from 2015 data (Pre Intervention)

```{r pairstable}
#Write out the new dataset with treatment allocation
knitr::kable(pairs2[order(pairs2$pair, pairs2$treat), c(1:2, 14, 9, 12, 10)]) #Order by pairs, with control pair office first
```

### Outcomes

The overall outcomes collected at the individual level within the period
covered by the data pull included:

 - `welcome` Whether the person had a welcome meeting
 - `uiexit` Whether the person stopped making unemployment claims for 3 weeks in a row.
 - `employed` Whether a different company reports wages for the person in a quarter after the quarter they applied for UI
 - `time.to.exit` The number of days between making the first claim and exiting the program.
 - `plans` Number of action plans uploaded by an individual in the program
 - `wages` The wages earned by the person (?average? ?weekly?)


```{r results=FALSE}
outcomes <- c("welcome","uiexit","employed","time.to.exit","plans","wages")
summary(or.df[,outcomes])
aggvars <- c("insert",outcomes)
```

```{r}
table(or.df$uiexit,months(or.df$claim.date))
```

## Employment Center Level Analysis

Because the intervention was randomized at the level of the employment office, we compare outcomes at that level: did offices in the new program display outcomes that differed from the offices operating under the status quo program? 

```{r}
# Remove Cases after Sept 30, 2016
or.df<-or.df[or.df$claim.date<=ymd("2016-09-30"),]
## This next creates more columns than I would like
officedat <- or.df[or.df$claim.date<=ymd("2016-09-30"),] %>% group_by(fonum) %>% summarize_at(aggvars,.funs=funs(mean(.,na.rm=TRUE),sum(is.na(.)),n()))
officedf <- as.data.frame(officedat[,c("fonum",grep("mean",names(officedat),value=TRUE),"wages_sum","insert_n")])
row.names(officedf)<-officedf$fonum
names(officedf)[names(officedf)%in%c("wages_sum","insert_n")] <- c("missingwage","N")


## wages_sum is the number of people with missing wages data

## Check that the calculation of the mean within office is correct by looking at one office
stopifnot(all.equal(with(or.df,mean(welcome[fonum=="210"])),officedf["210","welcome_mean"]))
```


```{r mergeontreatment}
## Add treatment assignment info
officedf$fonum <- as.numeric(officedf$fonum)
pairs2$fonum <- pairs2$fo.num
wrkdat <- inner_join(officedf,pairs2,by="fonum")
wrkdat <- wrkdat[order(wrkdat$pair,wrkdat$treat),]
row.names(wrkdat) <- as.character(wrkdat$fonum)

pairweights<-wrkdat %>% group_by(pair) %>% dplyr::summarise(
    weight=sum(N)/2 # weight by mean size of pair
 
)
n<-sum(wrkdat$N)
pairweights$weight<-pairweights$weight/n
wrkdat<-inner_join(wrkdat,pairweights,by="pair")



```


```{r}
## Pair-mean center to handle the pairing:

wrkdat$treatMD <- with(wrkdat,treat - ave(treat,pair))
for(i in outcomes){
	message(i)
	nm <- paste(i,"_mean",sep="")
	wrkdat[,paste(i,"MD",sep="")]<- wrkdat[[nm]] - ave(wrkdat[[nm]],wrkdat$pair)
}
knitr::kable(wrkdat) #If this table is too big it could be broken into pieces

## Check the pair centering
stopifnot(all.equal(wrkdat$welcomeMD[wrkdat$pair=="a"],
		    with(wrkdat[wrkdat$pair=="a",], welcome_mean - mean(welcome_mean))
		    ) )
```

# Office Level Effects

```{r estimates, results='hide'}
## First, just show two equivalent ways to get the ITT
# Note equal weights within pairs gives same values, unequal weights gives different values
welcomeATEa <- lm(welcome_mean~treat+pair,data=wrkdat,weights=weight)
welcomeATE <- lm(welcomeMD~treatMD,data=wrkdat,weights=weight)
summary(welcomeATE)
summary(welcomeATEa)



#stopifnot(all.equal(coef(welcomeATEa)[[2]],coef(welcomeATE)[[2]]))

## Next calc the diff of means for all outcomes
aggoutcomes <- paste(outcomes,"MD",sep="")
theITTS <- sapply(aggoutcomes,function(v){
			  coef(lm(reformulate("treatMD",response=v),data=wrkdat,weights=weight))[[2]]
})
theITTS
```


```{r ripvalues}
set.seed(20170324)

B <- gl(7,2)
N <- 14
K <- prod(rep(choose(2,1),N/2))
omega <- data.frame(matrix(-0.5,N,K))
blocksizes <- rep(2,7)
blocktreated <- rep(1,7)
trtedlocs<-mapply(function(x,m){ combn(x,m) },x=blocksizes,m=blocktreated,SIMPLIFY=FALSE)
trtedlocsMat <- t(expand.grid(trtedlocs))
trtloc_b <- lapply(split(trtedlocsMat,B),function(x){ cbind(x,1:128) })
om_b <- split(omega,B)
## Test
##Om_b[[1]][trtloc_b[[1]]]<-1
##Om_b[[1]]
for(i in 1:length(om_b)){
   om_b[[i]][trtloc_b[[i]]]<-.5
 }
 omega<-do.call("rbind",om_b)
 stopifnot(dim(unique(omega))==c(14,128))
 test1 <- apply(omega,2,function(x){ tapply(x,B,sum) })
 stopifnot(all(test1==0))

# # create dataframe with 128 possible paired assignements
# omega<-mosaic::do(1000)*mosaic::shuffle(wrkdat$treatMD,group=wrkdat$pair)
# omega<-data.frame(t(omega[!duplicated(omega),]))
# stopifnot(dim(omega)[2]==2^7)



# For each outcome generate null distribution, return pvalue
theps<-rep(NA,length(aggoutcomes))
names(theps)<-aggoutcomes
for(i in aggoutcomes){
    theNull<-sapply(1:128,function(x){
    coef(lm(reformulate("omega[,x]",response=i),data=wrkdat,weights=weight))[[2]]})
    #left<-mean(theNull<=theITTS[i])
    #right<-mean(theNull>=theITTS[i])
    #print(c(left,right))
    #theps[i]<-2*c(left,right)[which.min(c(left,right))]
    theps[i]<-mean(abs(theNull)>=abs(theITTS[i]))

}





```


```{r tests, eval=F,results='hide'}
# Commenting out old code,
# coin package doesn't seem to be happy with weights

## The coin package allows us to use permutations/repetitions of the experiment to assess the hypothesis of no effects in a block-randomized experiment
wrkdat$treatF <- factor(wrkdat$treat,levels=c("1","0"))
set.seed(20161102)
class(wrkdat$pair)
classwrkdat[,"treatF"]
test<-new("IndependenceProblem",x=data.frame(wrkdat[,"treatF"]),
          y=data.frame(wrkdat[,"welcome_mean"]),
          block=wrkdat$pair,
          weights=rep(5,14)
          )
test@weights<-wrkdat$weight
class(test@weights)
class(test)
oneway_test(test,distribution=approximate(B=5000))
welcomeTest <- oneway_test(welcome_mean ~ treatF | pair, data=wrkdat,distribution=approximate(B=5000))
welcomeTest@method

## The following uses a large sample approximation to the randomization distribution (Mostly just a check on the preceding)
xb1<-xBalance(treat ~ welcome_mean,strata=list(pair=~pair),data=wrkdat,report="all")
xb1$results
wrkdat$weight
## Now do it for all
theps <- sapply(aggoutcomes,function(v){
			 results <- oneway_test(reformulate("treatF | pair ",response=v,weights=weight),data=wrkdat,weights=weight,distribution=approximate(B=5000))
			 return(pvalue(results))
})
theps

xBalance(treat~welcome_mean+strata(pair),data=wrkdat,report="all") 

welcomeATE
```


We present differences of means weighted by pair size below --- they are an unbiased estimator of the average treatment effect. To assess the amount of information we have against the null of no effects, we present the results of hypothesis tests. Because the sample size is so small, we use a randomization-based approach. In short, the intervention appeared effective at increasing the number of plans completed, but did not appear influence any of the remaining outcomes.

```{r}
cbind(theITTS,theps)
```






```{r}
# Test the null of no effects for any outcome
# Not sure how to implement with weights yet

npcresults <- NPC(wrkdat,tr.var="treat",tr.label="1",y.vars=paste(outcomes,"_mean",sep=""),block.var="pair",alternative="two.sided",return.matrix=TRUE,FWE.adj=FALSE,print.steps=FALSE)
npcresults$p.values

# Here we see that the evidence against the null of no effects for any outcome has a p-value of roughly `r npcresults$p.values[["NPC"]]`. I also show the individual p-values here because they are exact (as compared to the approximate ones above). Here using the Student's T test statistic.
```



# Dealing with Missing Wage Data

`r table(or.df$employed,is.na(or.df$wages),useNA="ifany")[2,2]` participants are flagged as being employed but have no available wage data. The cases are relatively evenly distributed throughout the study, but slighty more common in the later months (`r round((1-sum(prop.table(table(month(or.df$claim.date[or.df$employed==1&is.na(or.df$wages)==T])))[8:10]))*100,2)` percent are for claims made before October).

Below we present the main results for employment status and compare them to results that exclude cases with mising wage data or treat those cases as instances of unemployment. In all cases the effect of treatment is marginally negative and non-significant.


```{r}

# 308 respondens coded as employed but missing wage data
table(or.df$employed,is.na(or.df$wages),useNA="ifany")

# Roughly evenly distributed through out the sample.
sum(prop.table(table(month(or.df$claim.date[or.df$employed==1&is.na(or.df$wages)==T])))[8:10])

# Exclude cases who said they were employed but had no wage data
officedat_emp <- or.df[!or.df$employed==1&is.na(or.df$wages)==T|
                           or.df$employed==1&!is.na(or.df$wages)==T
                       ,] %>% group_by(fonum) %>% summarize_at(aggvars,.funs=funs(mean(.,na.rm=TRUE),sum(is.na(.)),n()))
officedf_emp <- as.data.frame(officedat_emp[,c("fonum",grep("mean",names(officedat_emp),value=TRUE),"wages_sum","insert_n")])
row.names(officedf_emp)<-officedf_emp$fonum
names(officedf_emp)[names(officedf_emp)%in%c("wages_sum","insert_n")] <- c("missingwage","N")

officedf_emp$fonum <- as.numeric(officedf_emp$fonum)
pairs2$fonum <- pairs2$fo.num
wrkdat_emp <- inner_join(officedf_emp,pairs2,by="fonum")
wrkdat_emp <- wrkdat_emp[order(wrkdat_emp$pair,wrkdat_emp$treat),]
row.names(wrkdat_emp) <- as.character(wrkdat_emp$fonum)
wrkdat_emp$treatMD <- with(wrkdat_emp,treat - ave(treat,pair))
for(i in outcomes){
	message(i)
	nm <- paste(i,"_mean",sep="")
	wrkdat_emp[,paste(i,"MD",sep="")]<- wrkdat_emp[[nm]] - ave(wrkdat_emp[[nm]],wrkdat_emp$pair)
}
pairweights_emp<-wrkdat_emp %>% group_by(pair) %>% dplyr::summarise(
    weight=sum(N)/2
)
n<-sum(wrkdat_emp$N)
pairweights_emp$weight<-pairweights_emp$weight/n
wrkdat_emp<-inner_join(wrkdat_emp,pairweights_emp,by="pair")


employedATE <- coef(lm(employedMD~treatMD,data=wrkdat_emp,weights=weight))[2]
wrkdat_emp$treatF <- factor(wrkdat_emp$treat,levels=c("1","0"))
employedATE_pval<-pvalue(oneway_test(employed_mean~treatF|pair,data=wrkdat_emp,distribution=approximate(B=5000)))

# Create alternate employment flag
or.df$employed_alt<-or.df$employed
# Treat missing wage cases as unemployed
or.df$employed_alt[or.df$employed==1&is.na(or.df$wages)==T]<-0

officedat_emp2 <- or.df %>% group_by(fonum) %>% summarize_at(c(aggvars,"employed_alt"),.funs=funs(mean(.,na.rm=TRUE),sum(is.na(.)),n()))
officedf_emp2 <- as.data.frame(officedat_emp2[,c("fonum",grep("mean",names(officedat_emp2),value=TRUE),"wages_sum","insert_n")])
row.names(officedf_emp2)<-officedf_emp2$fonum
names(officedf_emp2)[names(officedf_emp2)%in%c("wages_sum","insert_n")] <- c("missingwage","N")

officedf_emp2$fonum <- as.numeric(officedf_emp2$fonum)
pairs2$fonum <- pairs2$fo.num
wrkdat_emp2 <- inner_join(officedf_emp2,pairs2,by="fonum")
wrkdat_emp2 <- wrkdat_emp2[order(wrkdat_emp2$pair,wrkdat_emp2$treat),]
row.names(wrkdat_emp2) <- as.character(wrkdat_emp2$fonum)
wrkdat_emp2$treatMD <- with(wrkdat_emp2,treat - ave(treat,pair))
for(i in c(outcomes,"employed_alt")){
	message(i)
	nm <- paste(i,"_mean",sep="")
	wrkdat_emp2[,paste(i,"MD",sep="")]<- wrkdat_emp2[[nm]] - ave(wrkdat_emp2[[nm]],wrkdat_emp2$pair)
}
pairweights_emp2<-wrkdat_emp2 %>% group_by(pair) %>% dplyr::summarise(
    weight=sum(N)/2
)
n<-sum(wrkdat_emp2$N)
pairweights_emp2$weight<-pairweights_emp2$weight/n
wrkdat_emp2<-inner_join(wrkdat_emp2,pairweights_emp2,by="pair")

employedATEalt <- coef(lm(employed_altMD~treatMD,data=wrkdat_emp2,weights=weight))[2]
wrkdat_emp2$treatF <- factor(wrkdat_emp2$treat,levels=c("1","0"))
employedATE_pval_alt<-pvalue(oneway_test(employed_alt_mean~treatF|pair,data=wrkdat_emp2,distribution=approximate(B=5000)))


employed_tab<-cbind(rbind(theITTS[3],employedATE,employedATEalt),rbind(theps[3],employedATE_pval,employedATE_pval_alt))
rownames(employed_tab)<-c("Main Result","Removing Cases with Missing Wages","Recoding Missing Wages as Unemployed")
colnames(employed_tab)<-c("ITT","p-value")
```

```{r,results="markdown"}
kable(employed_tab,caption = "ITT for Treatment on Employment Status")
```


# Effects Excluding Those With a Short Time on the Program

```{r,negativeexit,echo=F,results="hide"}
#sum(or.df$time.to.exit<=0)/dim(or.df)[1]
#sum(sort(prop.table(with(or.df[or.df$time.to.exit<=0,],table(time.to.exit))),decreasing = T)[1:5])

officedat_noneg <- or.df[or.df$time.to.exit>=0,] %>% group_by(fonum) %>% summarize_at(aggvars,.funs=funs(mean(.,na.rm=TRUE),sum(is.na(.)),n()))
officedf_noneg <- as.data.frame(officedat_noneg[,c("fonum",grep("mean",names(officedat_noneg),value=TRUE),"wages_sum","insert_n")])
row.names(officedf_noneg)<-officedf_noneg$fonum
names(officedf_noneg)[names(officedf_noneg)%in%c("wages_sum","insert_n")] <- c("missingwage","N")

## Add treatment assignment info
officedf_noneg$fonum <- as.numeric(officedf_noneg$fonum)
pairs2$fonum <- pairs2$fo.num
wrkdat_noneg <- inner_join(officedf_noneg,pairs2,by="fonum")
wrkdat_noneg <- wrkdat_noneg[order(wrkdat_noneg$pair,wrkdat_noneg$treat),]
row.names(wrkdat_noneg) <- as.character(wrkdat_noneg$fonum)

## Pair-mean center to handle the pairing:

wrkdat_noneg$treatMD <- with(wrkdat_noneg,treat - ave(treat,pair))
for(i in outcomes){
	message(i)
	nm <- paste(i,"_mean",sep="")
	wrkdat_noneg[,paste(i,"MD",sep="")]<- wrkdat_noneg[[nm]] - ave(wrkdat_noneg[[nm]],wrkdat_noneg$pair)
}

wrkdat_noneg$treatMD <- with(wrkdat_noneg,treat - ave(treat,pair))
for(i in outcomes){
	message(i)
	nm <- paste(i,"_mean",sep="")
	wrkdat_noneg[,paste(i,"MD",sep="")]<- wrkdat_noneg[[nm]] - ave(wrkdat_noneg[[nm]],wrkdat_noneg$pair)
}
pairweights_noneg<-wrkdat_noneg %>% group_by(pair) %>% dplyr::summarise(
    weight=sum(N)/2
)
n<-sum(wrkdat_noneg$N)
pairweights_noneg$weight<-pairweights_noneg$weight/n
wrkdat_noneg<-inner_join(wrkdat_noneg,pairweights_noneg,by="pair")


aggoutcomes <- paste(outcomes,"MD",sep="")
theITTS_noneg <- sapply(aggoutcomes,function(v){
			  coef(lm(reformulate("treatMD",response=v),data=wrkdat_noneg,weights=weight))[[2]]
})

theps_noneg<-rep(NA,length(aggoutcomes))
names(theps_noneg)<-aggoutcomes
for(i in aggoutcomes){
    theNull<-sapply(1:128,function(x){
    coef(lm(reformulate("omega[,x]",response=i),data=wrkdat_noneg,weights=weight))[[2]]})
    #left<-mean(theNull<=theITTS[i])
    #right<-mean(theNull>=theITTS[i])
    #print(c(left,right))
    #theps_noneg[i]<-2*c(left,right)[which.min(c(left,right))]
    theps_noneg[i]<-mean(abs(theNull)>=abs(theITTS[i]))

}

#wrkdat_noneg$treatF <- factor(wrkdat_noneg$treat,levels=c("1","0"))
#

# theps_noneg <- sapply(aggoutcomes,function(v){
# 			 results <- oneway_test(reformulate("treatF | pair ",response=v),data=wrkdat_noneg,distribution=approximate(B=5000))
# 			 return(pvalue(results))
# })
# 
# npcresults_nonneg <- NPC(wrkdat_noneg,tr.var="treat",tr.label="1",y.vars=paste(outcomes[-5],"_mean",sep=""),block.var="pair",alternative="two.sided",return.matrix=TRUE,FWE.adj=FALSE,print.steps=FALSE)

```


A small portion of the cases appear to have "negative" times to exiting the program, (N=`r sum(or.df$time.to.exit<=0)` or `r round(sum(or.df$time.to.exit<=0)/dim(or.df)[1]*100,1)` percent of cases)

Time to exiting the program is calculated by the subtracting participant's initial unemployment claim date from the date they exited the unemployment insurance. An individual is considered to have exited the program after three weeks of not claiming benefits. Thus exit dates are always on Sundays when a new claim period begins, while initial claim date may be any day during the week. In fact nearly `r round(sum(sort(prop.table(with(or.df[or.df$time.to.exit<=0,],table(time.to.exit))),decreasing = T)[1:5])*100,1)` percent of these cases have values implying exit dates within seven days of the claim process, suggesting most of the negative cases are due to disconnect between how claims are processed and exit times are calculated. Still, as a robustness check, we repeat the main analysis, removing participants with negative times to exit. The results are essentially unchanged.


```{r}
cbind(theITTS_noneg,theps_noneg)
```



# Effects on Those With More Time in the Program?

We might expect that the effects may be larger among individuals who have spent more time in the program. In particular since wages are reported quarterly, we now re-do the analysis on the subset of respondents with initial claims dates before June 30, 2016,  (N=`r dim(or.df[or.df$claim.date<=ymd("2016-06-30"),] )[1]`


```{r agg2}

#dim(or.df[or.df$claim.date<=ymd("2016-06-30"),] )[1]
## This next creates more columns than I would like
officedat2 <- or.df[or.df$claim.date<=ymd("2016-06-30"),] %>% group_by(fonum) %>% summarize_at(aggvars,.funs=funs(mean(.,na.rm=TRUE),sum(is.na(.)),n()))
officedf2 <- as.data.frame(officedat2[,c("fonum",grep("mean",names(officedat2),value=TRUE),"wages_sum","insert_n")])
row.names(officedf2)<-officedf2$fonum
names(officedf2)[names(officedf2)%in%c("wages_sum","insert_n")] <- c("missingwage","N")

## Add treatment assignment info
officedf2$fonum <- as.numeric(officedf2$fonum)
pairs2$fonum <- pairs2$fo.num
wrkdat2 <- inner_join(officedf2,pairs2,by="fonum")
wrkdat2 <- wrkdat2[order(wrkdat2$pair,wrkdat2$treat),]
row.names(wrkdat2) <- as.character(wrkdat2$fonum)
```


```{r}
## Pair-mean center to handle the pairing:

wrkdat2$treatMD <- with(wrkdat2,treat - ave(treat,pair))
for(i in outcomes){
	message(i)
	nm <- paste(i,"_mean",sep="")
	wrkdat2[,paste(i,"MD",sep="")]<- wrkdat2[[nm]] - ave(wrkdat2[[nm]],wrkdat2$pair)
}
knitr::kable(wrkdat2) #If this table is too big it could be broken into pieces

## Check the pair centering
stopifnot(all.equal(wrkdat2$welcomeMD[wrkdat2$pair=="a"],
		    with(wrkdat2[wrkdat2$pair=="a",], welcome_mean - mean(welcome_mean))
		    ) )

pairweights2<-wrkdat2 %>% group_by(pair) %>% dplyr::summarise(
    weight=sum(N)/2
)
n<-sum(wrkdat2$N)
pairweights2$weight<-pairweights2$weight/n
wrkdat2<-inner_join(wrkdat2,pairweights2,by="pair")

```



```{r estimates2, results='hide'}
theITTS2 <- sapply(aggoutcomes,function(v){
			  coef(lm(reformulate("treatMD",response=v),data=wrkdat2,weights=weight))[[2]]
})
theITTS2
```


```{r tests2, results='hide'}
theps2<-rep(NA,length(aggoutcomes))
names(theps2)<-aggoutcomes
for(i in aggoutcomes){
    theNull<-sapply(1:128,function(x){
    coef(lm(reformulate("omega[,x]",response=i),data=wrkdat2,weights=weight))[[2]]})
    #left<-mean(theNull<=theITTS[i])
    #right<-mean(theNull>=theITTS[i])
    #print(c(left,right))
    #theps2[i]<-2*c(left,right)[which.min(c(left,right))]
    theps2[i]<-mean(abs(theNull)>=abs(theITTS[i]))

}

## The coin package allows us to use permutations/repetitions of the experiment to assess the hypothesis of no effects in a block-randomized experiment
# wrkdat2$treatF <- factor(wrkdat2$treat,levels=c("1","0"))
# set.seed(20161102)
# 
# ## Now do it for all
# theps2 <- sapply(aggoutcomes,function(v){
# 			 results <- oneway_test(reformulate("treatF | pair ",response=v),data=wrkdat2,distribution=approximate(B=5000))
# 			 return(pvalue(results))
# })
# theps2
```

Here are the estimated ITTs and p-values for the null of no effects among the group that had the most time in the program.

```{r}
cbind(theITTS2,theps2)
```


```{r,eval=F}
# Test the null of no effects for any outcome

npcresults2 <- NPC(wrkdat2,tr.var="treat",tr.label="1",y.vars=paste(outcomes[-5],"_mean",sep=""),block.var="pair",alternative="two.sided",return.matrix=TRUE,FWE.adj=FALSE,print.steps=FALSE)
npcresults2$p.values

# Here we see that the evidence against the null of no effects for any outcome with a p-value of roughly `r npcresults2$p.values[["NPC"]]`. I also show the individual p-values here because they are exact (as compared to the approximate ones above). The test statistic used here, by default, is Student's T.
```

## Effects conditional on completing the welcome

```{r}
officedat_welcome <- or.df[or.df$welcome==1,] %>% group_by(fonum) %>% summarize_at(aggvars,.funs=funs(mean(.,na.rm=TRUE),sum(is.na(.)),n()))
officedf_welcome <- as.data.frame(officedat_welcome[,c("fonum",grep("mean",names(officedat_welcome),value=TRUE),"wages_sum","insert_n")])
row.names(officedf_welcome)<-officedf_welcome$fonum
names(officedf_welcome)[names(officedf_welcome)%in%c("wages_sum","insert_n")] <- c("missingwage","N")

## Add treatment assignment info
officedf_welcome$fonum <- as.numeric(officedf_welcome$fonum)
pairs2$fonum <- pairs2$fo.num
wrkdat_welcome <- inner_join(officedf_welcome,pairs2,by="fonum")
wrkdat_welcome <- wrkdat_welcome[order(wrkdat_welcome$pair,wrkdat_welcome$treat),]
row.names(wrkdat_welcome) <- as.character(wrkdat_welcome$fonum)

## Pair-mean center to handle the pairing:

wrkdat_welcome$treatMD <- with(wrkdat_welcome,treat - ave(treat,pair))
for(i in outcomes){
	message(i)
	nm <- paste(i,"_mean",sep="")
	wrkdat_welcome[,paste(i,"MD",sep="")]<- wrkdat_welcome[[nm]] - ave(wrkdat_welcome[[nm]],wrkdat_welcome$pair)
}

wrkdat_welcome$treatMD <- with(wrkdat_welcome,treat - ave(treat,pair))
for(i in outcomes){
	message(i)
	nm <- paste(i,"_mean",sep="")
	wrkdat_welcome[,paste(i,"MD",sep="")]<- wrkdat_welcome[[nm]] - ave(wrkdat_welcome[[nm]],wrkdat_welcome$pair)
}
pairweights_welcome<-wrkdat_welcome %>% group_by(pair) %>% dplyr::summarise(
    weight=sum(N)/2
)
n<-sum(wrkdat_welcome$N)
pairweights_welcome$weight<-pairweights_welcome$weight/n
wrkdat_welcome<-inner_join(wrkdat_welcome,pairweights_welcome,by="pair")


aggoutcomes <- paste(outcomes,"MD",sep="")
aggoutcomes<-aggoutcomes[-1] #Remove welcome
theITTS_welcome <- sapply(aggoutcomes,function(v){
			  coef(lm(reformulate("treatMD",response=v),data=wrkdat_welcome,weights=weight))[[2]]
})

theps_welcome<-rep(NA,length(aggoutcomes))
names(theps_welcome)<-aggoutcomes
for(i in aggoutcomes){
    theNull<-sapply(1:128,function(x){
    coef(lm(reformulate("omega[,x]",response=i),data=wrkdat_welcome,weights=weight))[[2]]})
    #left<-mean(theNull<=theITTS[i])
    #right<-mean(theNull>=theITTS[i])
    #print(c(left,right))
    #theps_welcome[i]<-2*c(left,right)[which.min(c(left,right))]
    theps_welcome[i]<-mean(abs(theNull)>=abs(theITTS[i]))

}
```


Here are the estimated ITTs and p-values for the null of no effects among the group that completed the initial welcome meeting.

```{r}
cbind(theITTS_welcome,theps_welcome)
```


## Separate Effects by Month

Below I show the estimated effects by month for each of the six outcomes.

```{r monthly}
# Wrapper Function to calculate by month

monthly_itts<-function(the_month=3){
    ## This next creates more columns than I would like
officedat_month <- or.df[month(or.df$claim.date)==the_month,] %>% group_by(fonum) %>% summarize_at(aggvars,.funs=funs(mean(.,na.rm=TRUE),sum(is.na(.)),n()))
officedf_month <- as.data.frame(officedat_month[,c("fonum",grep("mean",names(officedat_month),value=TRUE),"wages_sum","insert_n")])
row.names(officedf_month)<-officedf_month$fonum
names(officedf_month)[names(officedf_month)%in%c("wages_sum","insert_n")] <- c("missingwage","N")

## Add treatment assignment info
officedf_month$fonum <- as.numeric(officedf_month$fonum)
pairs2$fonum <- pairs2$fo.num
wrkdat_month <- inner_join(officedf_month,pairs2,by="fonum")
wrkdat_month <- wrkdat_month[order(wrkdat_month$pair,wrkdat_month$treat),]
row.names(wrkdat_month) <- as.character(wrkdat_month$fonum)

## Pair-mean center to handle the pairing:

wrkdat_month$treatMD <- with(wrkdat_month,treat - ave(treat,pair))
for(i in outcomes){
	message(i)
	nm <- paste(i,"_mean",sep="")
	wrkdat_month[,paste(i,"MD",sep="")]<- wrkdat_month[[nm]] - ave(wrkdat_month[[nm]],wrkdat_month$pair)
}

wrkdat_month$treatMD <- with(wrkdat_month,treat - ave(treat,pair))
for(i in outcomes){
	message(i)
	nm <- paste(i,"_mean",sep="")
	wrkdat_month[,paste(i,"MD",sep="")]<- wrkdat_month[[nm]] - ave(wrkdat_month[[nm]],wrkdat_month$pair)
}
pairweights_month<-wrkdat_month %>% group_by(pair) %>% dplyr::summarise(
    pairN=sum(N)
)
pairweights_month<-wrkdat_month %>% group_by(pair) %>% dplyr::summarise(
    weight=sum(N)/2
)
n<-sum(wrkdat_month$N)
pairweights_month$weight<-pairweights_month$weight/n
wrkdat_month<-inner_join(wrkdat_month,pairweights_month,by="pair")
if(the_month>9){outcomes<-outcomes[-6]} # no wage data after september
aggoutcomes <- paste(outcomes,"MD",sep="")
theITTS_month <- sapply(aggoutcomes,function(v){
			  coef(lm(reformulate("treatMD",response=v),data=wrkdat_month,weights=weight))[[2]]
})

theps_month<-rep(NA,length(aggoutcomes))
names(theps_month)<-aggoutcomes
for(i in aggoutcomes){
    theNull<-sapply(1:128,function(x){
    coef(lm(reformulate("omega[,x]",response=i),data=wrkdat_month,weights=weight))[[2]]})
    #left<-mean(theNull<=theITTS[i])
    #right<-mean(theNull>=theITTS[i])
    #print(c(left,right))
    #theps_month[i]<-2*c(left,right)[which.min(c(left,right))]
    theps_month[i]<-mean(abs(theNull)>=abs(theITTS[i]))

}
# wrkdat_month$treatF <- factor(wrkdat_month$treat,levels=c("1","0"))

# theps_month <- sapply(aggoutcomes,function(v){
# 			 results <- oneway_test(reformulate("treatF | pair ",response=v),data=wrkdat_month,distribution=approximate(B=5000))
# 			 return(pvalue(results))
# })

return(data.frame(theITTS_month,theps_month,the_month,outcomes))
}

march_itt<-monthly_itts(the_month=3)
april_itt<-monthly_itts(the_month=4)
may_itt<-monthly_itts(the_month=5)
june_itt<-monthly_itts(the_month=6)
july_itt<-monthly_itts(the_month=7)
august_itt<-monthly_itts(the_month=8)
september_itt<-monthly_itts(the_month=9)
#october_itt<-monthly_itts(the_month=10)
#november_itt<-monthly_itts(the_month=11)
#december_itt<-monthly_itts(the_month=12)

monthly_df<-rbind(march_itt,april_itt,may_itt,june_itt,july_itt,august_itt,september_itt)
monthly_df$outcomes<-factor(monthly_df$outcomes)
library(ggplot2)       
monthly_plot<-ggplot(monthly_df,aes(x=the_month,y=theITTS_month))+geom_line()+facet_wrap(~outcomes,scales = "free")+labs(x="Month",y="ITT")


```

```{r,results="markup"}
monthly_plot
```


