---
title: "Prelim Analysis/Data Inspection"
author: "Daniel Shephard, Jake Bowers, Michael DiDomenico"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: html_document
---

```{r}
load("Data/or.df.TOTAL.rda")
```


```{r Welcomes_by_Office, echo=FALSE}

#Create a table showing the number of observations by month for each office
offices <- c(unique(or.df.TOTAL$'277.Field.Office'))[1:14]
months <- c('mar', 'apr', 'may', 'jun', 'jul', 'aug', 'sep')
desc.table.1 <- matrix(nrow = 14, ncol = 7,
		       dimnames = list(offices, months))

#Add the counts of welcomes for March
for(i in 1:14)
{
	desc.table.1[i, 1] <- length(which(or.df1$`277.Field.Office`==offices[i]))
}

#Add the counts of welcomes for April
for(i in 1:14)
{
	desc.table.1[i, 2] <- length(which(or.df2$`277.Field.Office`==offices[i]))
}

#Add the counts of welcomes for May
for(i in 1:14)
{
	desc.table.1[i, 3] <- length(which(or.df3$`277.Field.Office`==offices[i]))
}

#Add the counts of welcomes for June
for(i in 1:14)
{
	desc.table.1[i, 4] <- length(which(or.df4$`277.Field.Office`==offices[i]))
}

#Add the counts of welcomes for July
for(i in 1:14)
{
	desc.table.1[i, 5] <- length(which(or.df5$`277.Field.Office`==offices[i]))
}

#Add the counts of welcomes for August
for(i in 1:14)
{
	desc.table.1[i, 6] <- length(which(or.df6$`277.Field.Office`==offices[i]))
}

#Add the counts of welcomes for September
for(i in 1:14)
{
	desc.table.1[i, 7] <- length(which(or.df7$`277.Field.Office`==offices[i]))
}

knitr::kable(desc.table.1)
```

##Table of welcome conversations by field office
The table above shows the number of welcome conversations conducted at each of the field offices in each month.


#Next Model the Outcome of UI Exit w/o Treatment

First, we deterimined which variables were most predictive of the outcome of interest (exiting UI) in order to remove unnecessary noise to amplify the power of the study.

The variables that were most predictive were **Worker Profile Score**, **Education Level**, and **Disability Status**.

```{r Model_Creation}
#Create a function to predict UI Exit from a given "Variable"
dsMonthlyExitModel <- function(df, variable)
{

	df$ui.exit.rc <- 0
	df[df$Exited.UI.Flag=="Y", "ui.exit.rc"] <- 1
	df$Worker.Profile.Score <- as.numeric(df$Worker.Profile.Score)

	tmp.m1 <- lm(ui.exit.rc ~ df[, variable], data = df)
	print(variable)
	summary(tmp.m1)
	return(tmp.m1)
}

#Create a function to run a model predicting UI Exit from 8 variables
dsMonthlyExitModel2 <- function(df, variable)
{

	df$ui.exit.rc <- 0
	df[df$Exited.UI.Flag=="Y", "ui.exit.rc"] <- 1
	df$Worker.Profile.Score <- as.numeric(df$Worker.Profile.Score)

	tmp.m1 <- lm(ui.exit.rc ~ df[, variable[1]] +
		     df[, variable[2]] +
		     df[, variable[3]] +
		     df[, variable[4]] +
		     df[, variable[5]] +
		     df[, variable[6]] +
		     df[, variable[7]] +
		     df[, variable[8]], data = df)
	print(variable)
	summary(tmp.m1)
	return(tmp.m1)
}

#Create a model to predict UI Exit from 3 variables -- the variables that were predictive in the 8 variable model
dsMonthlyExitModel3 <- function(df, variable)
{

	df$ui.exit.rc <- 0
	df[df$Exited.UI.Flag=="Y", "ui.exit.rc"] <- 1
	df$Worker.Profile.Score <- as.numeric(df$Worker.Profile.Score)

	tmp.m1 <- lm(ui.exit.rc ~ df[, variable[1]] +
		     df[, variable[2]] +
		     df[, variable[3]], data = df)
	print(variable)
	summary(tmp.m1)
	return(tmp.m1)
}


#Use March and April Data first to determine if the 3 variables are significantly predictive
or.df1$ui.exit.rc <- 0
or.df1[or.df1$Exited.UI.Flag=="Y", "ui.exit.rc"] <- 1
or.df1$Worker.Profile.Score <- as.numeric(or.df1$Worker.Profile.Score)
variableList <- c("WBA", "Worker.Profile.Score", "Age", "Education.Level", "Race.White.Flag", "Ethnicity.Hispanic.Latino.Flag", "Ex.Offender.Flag", "Disability")
stats <- c("Estimate", "sigma", "tstat", "pvalue")

modelsAPRIL <- matrix(ncol = 4, nrow = 8, dimnames = list(variableList, stats))
for(i in 1:8)
{
	modelsAPRIL[i, ] <- summary(dsMonthlyExitModel(or.df2, variableList[i]))$coefficient[2, ]
}
knitr::kable(modelsAPRIL) #These are the esimates of the first level of the models for APRIL (because March only had one week of PEP)

#Make a multivariate model with all 8 predictive variables using both March and April
comb.m1 <- (dsMonthlyExitModel2(or.df1, variableList)) #March
comb.m2 <- (dsMonthlyExitModel2(or.df2, variableList)) #April
summary(comb.m1)
summary(comb.m2)

#Use this model going forward, Worker.Profile.Score + Education.Level + Disability
comb.m3 <- dsMonthlyExitModel3(or.df2, variableList[c(2, 4, 8)]) #April
summary(comb.m3) #April

#Run the models for May to see if it holds true -- Yes for Profile and Disability, no for Education
comb.m3.MAY <- dsMonthlyExitModel3(or.df3, variableList[c(2, 4, 8)])
summary(comb.m3.MAY)

```


#Pair Offices
Now that the variables to be used for the covariate adjustment have been determined, create pair centered variables. 

Pairs created by both urban, geography, size (number of people who came for UI), and quality of staff (judgement of the management in Oregon). Pairs provided by the management of the Oregon State program.

```{r Pull_Pairs}
#----ORIGINAL CODE FOR RANDOMIZATION------
pairfilepath <- "C:/Users/DanielDShephard/Documents/SBST/DOL/UI Oregon/Analysis/OR_Analysis/Oregon_Pairs_14-Field-Offices_20160111.csv" 
pairs2 <- read.csv(pairfilepath, row.names = 1)
pairs2 <- pairs2[order(pairs2$pair), ] #ensure that all pairs are next to each other
#Random generation of uniform distribution x14, one number per field office
#The highest number in the pair is allocated to the treatment group
pairs2$treat <- 0
set.seed(20160108) #note, ran once with 20150108, but changed to match date
ran.no <- runif(14)
for (i in 1:7)
{
	if(ran.no[i*2] > ran.no[(i*2)-1])
	{
		pairs2[i*2, "treat"] <- 1
	}
	else
	{
		pairs2[i*2-1, "treat"] <- 1
	}

}

#Check the frequency of assignment with different seeds to ensure 50% chance
#Run allocation logic used above through 1000 different seeds
pair2 <- NA
pair2 <- as.numeric(pair2)
for (q in 1:1000) #run 1000 simulations of different seeds
{
	set.seed(20160107+q)
	ran.no <- runif(14)
	for (i in 1:7) #for a set of 7 pairs with the current seed do the following
	{
		if(ran.no[i*2] > ran.no[(i*2)-1]) #if the even number in the pair is greater than the odd number
		{
			pair2 <- rbind(pair2, 1) #then give it a treatment value of 1
		}
		else
		{
			pair2 <- rbind(pair2, 0) #else give the second in the pair a value of 0
		}
	}
}
treat.probability <- mean(pair2, na.rm=TRUE)
samp.m1 <- (lm(ui.wks ~ treat, data=pairs2)) #original scores
samp.m2 <- (lm(tot ~ treat, data=pairs2)) #original scores
summary(samp.m1)
summary(samp.m2)
#Write out the new dataset with treatment allocation
#The number of welcomes and average weeks on UI is from 2015 data (Pre Intervention)
knitr::kable(pairs2[order(pairs2$pair, pairs2$treat), c(1:2, 14, 9, 12, 10)]) #Order by pairs, with control pair office first

```


```{r Pair_Centering}
#----Missing Values so COV adjust in last step now-----
#STEP 1: create covariate adjusted UI exit residuals
#or.dfa <- ds.na.omit(or.df2)
#m1.res <- ds.cov.uiexit(or.dfa)
#summary(m1.res) #check model
#or.dfa$uiexit.resid <- m1.res$residuals #insert residuals into dataframe to be the main outcome

#This is now the first step
#STEP 2: Create pair centered data
ds.create.pairs <- function(df)
{
	df$pair <- as.character(NA)
	df$treat <- 0
	for(i in 1:14)
	{
		df[as.character(df$fo)==pairs2[i, "fo.num"], "pair"] <- as.character(pairs2[i, "pair"])
		df[as.character(df$fo)==pairs2[i, "fo.num"], "treat"] <- pairs2[i, "treat"]
	}

	return(df)
}

or.dfb <- or.df.TOTAL
or.dfb$pair <- as.character(NA)
or.dfb$treat <- 0
for(i in 1:14)
{
	or.dfb[which(as.character(or.dfb$`277.Field.Office`)==pairs2[i, "fo.num"]), "pair"] <- as.character(pairs2[i, "pair"])
	or.dfb[which(as.character(or.dfb$`277.Field.Office`)==pairs2[i, "fo.num"]), "treat"] <- pairs2[i, "treat"]
}

tmp.pair <- c("a", "b", "c", "d", "e", "f", "g")

or.dfb$uiexit.pair.delta <- NA #create a pair variable
or.dfb$employed.pair.delta <- NA

for(i in 1:7)
{
	pair.mean <- mean(or.dfb[or.dfb$pair==tmp.pair[i], "uiexit"], na.rm=TRUE)
	print(pair.mean)
	or.dfb[which(or.dfb$pair==tmp.pair[i]), "uiexit.pair.delta"] <-   or.dfb[which(or.dfb$pair==tmp.pair[i]),"uiexit"] - pair.mean #create pair centered variable

	pair.mean.emp <- mean(or.dfb[or.dfb$pair==tmp.pair[i], "employed"], na.rm = TRUE)
	print(pair.mean.emp)
	or.dfb[which(or.dfb$pair==tmp.pair[i]), "employed.pair.delta"] <- or.dfb[which(or.dfb$pair==tmp.pair[i]), "employed"] - pair.mean.emp
}


```

#Main Effects
Now run the analysis to see if there is a treatment effect clustering by office

```{r Main_Effects}

library(nlme)
library(lmtest)
library(sandwich)
library(car)
or.dfb$fo <- as.factor(or.dfb$`277.Field.Office`)
or.dfb <- or.dfb[which(!is.na(or.dfb$PEP.Unique.ID)), ]

#STEP 3: run main model regressed on treatment and controlling for office cluster

#Question 1: Did the insert result in more completed Welcomes?
m.insert.welcome <- lme(welcome ~ insert,
			random = ~1 | fo,
			data = or.dfb)
summary(m.insert.welcome)

#Question 2a: ITT Did being assigned to a treatment FO increase UI exit rates, non covariate adjusted
m.exit.itt1 <- lme(uiexit ~ treat,
		   random = ~1 | fo,
		   data = or.dfb)
summary(m.exit.itt1)

m.exit.itt1b <- lme(uiexit.pair.delta ~ treat,
		    random = ~1 | fo,
		    data = or.dfb)
summary(m.exit.itt1b)

#Question 2b: ITT Did being assigned to a treatment FO increase UI exit rates, pair & covariate adjusted
#Create DF dropping missing cases
or.dfb$BYE <- as.numeric(or.dfb$BYE)
or.dfb.omit <- na.omit(or.dfb[c(1, 2, 5, 9, 30, 35, 42, 45, 46:56)])
m.exit.itt2 <- lme(uiexit.pair.delta ~ treat + BYE + Education.Level + Worker.Profile.Score +
		   Disability,
	   random = ~1 | fo,
	   data = or.dfb.omit)
summary(m.exit.itt2) #only complete cases with covariate adjustment on pair centered outcome

#Question 2c: TOT Among those that came in for a welcome meeting did PEP increase UI exit rates
m.exit.tot1 <- lme(uiexit ~ treat,
		   random = ~1 | fo,
		   data = or.dfb[or.dfb$welcome==1, ])
summary(m.exit.tot1)

#Question 2d: TOT Among those that came in for a welcome did PEP increase UI exit rates, covariate adjusted
m.exit.tot2 <- lme(uiexit.pair.delta ~ treat + BYE +  Worker.Profile.Score +
		   Disability,
	   random = ~1 | fo,
	   data = or.dfb.omit[or.dfb.omit$welcome==1, ]) #dropped Education to improve model convergence
summary(m.exit.tot2)

#Question 3: TOT 2.0 Among those that came for a welcome, did uploading a plan
or.dfb.omit2 <- na.omit(or.dfb[c(1, 2, 5, 30, 35, 42, 45, 46:56)])
or.dfb.omit2$plans <- as.numeric(or.dfb.omit2$`#.of.PEPP.Plans`)
or.dfb.omit2$emails <- as.numeric(or.dfb.omit2$`#.of.PEPP.Messages.Sent`)
m.exit.plans.tot3 <- lme(uiexit.pair.delta ~ treat*plans + BYE +  Worker.Profile.Score +
			 Disability,
		 random = ~1 | fo,
		 data = or.dfb.omit2[or.dfb.omit2$welcome==1, ]) #dropped Education to improve model
summary(m.exit.plans.tot3)

#----Only Individuals w/ BYE of 201718 or less [25th percentile]----
#This analysis is to see what the impact is when there is more time for follow-up
Hmisc::describe(as.numeric(or.dfb$BYE))
or.df.LT <- or.dfb[or.dfb$BYE<=201718,]#longer term follow-up group
or.df.LT$plans <- as.numeric(or.df.LT$`#.of.PEPP.Plans`)
or.df.LT.omit <- or.dfb.omit2[or.dfb.omit2$BYE<=201718, ]

m.exit.itt1.LT <- lme(uiexit ~ treat,
		      random = ~1 | fo,
		      data = or.df.LT)
summary(m.exit.itt1.LT)

#Absoluate & Relative Change

ds.confidence.table <- function(m, outcome)
{
	ppt <- summary(m)$coefficients$fixed["treat"]
	#relative <- summary(m)$coefficients$fixed["treat"] / summary(m)$coefficients$fixed["(Intercept)"]
	temp.ci <- (sqrt(diag(vcov(m))["treat"])*1.96)
	lower.ci <- ppt - temp.ci
	upper.ci <- ppt + temp.ci

	return(data.frame(outcome, ppt, lower.ci, upper.ci))
}

ds.confidence.table.plans <- function(m, outcome)
{
	ppt <- summary(m)$coefficients$fixed["plans"]
	#relative <- summary(m)$coefficients$fixed["plans"] / summary(m)$coefficients$fixed["(Intercept)"]
	temp.ci <- (sqrt(diag(vcov(m))["plans"])*1.96)
	lower.ci <- ppt - temp.ci
	upper.ci <- ppt + temp.ci

	return(data.frame(outcome, ppt, lower.ci, upper.ci))
}



```

#Employment Main Outcomes

```{r Main_Employment}
#----Employed Outcome----------
#Question 4a: ITT Did being assigned to a treatment FO increase UI employment, non covariate adjusted
m.employed.itt1 <- lme(employed~ treat,
		       random = ~1 | fo,
		       data = or.dfb)
summary(m.employed.itt1)

m.employed.itt1b <- lme(employed.pair.delta~ treat,
			random = ~1 | fo,
			data = or.dfb)
summary(m.employed.itt1b)

#Question 4b: ITT Did being assigned to a treatment FO increase employment rates, pair & covariate adjusted
#Create DF dropping missing cases
m.employed.itt2 <- lme(employed.pair.delta ~ treat + as.numeric(BYE) + Education.Level + Worker.Profile.Score +
		       Disability,
	       random = ~1 | fo,
	       data = or.dfb.omit)
summary(m.employed.itt2) #only complete cases with covariate adjustment on pair centered outcome

#Question 4c: TOT Among those that came in for a welcome meeting did PEP increase employment rates
m.employed.tot1 <- lme(employed~ treat,
		       random = ~1 | fo,
		       data = or.dfb[or.dfb$welcome==1, ])
summary(m.employed.tot1)

#Question 4d: TOT Among those that came in for a welcome did PEP increase employment, covariate adjusted
m.employed.tot2 <- lme(employed.pair.delta ~ treat + BYE +  Worker.Profile.Score +
		       Disability,
	       random = ~1 | fo,
	       data = or.dfb.omit[or.dfb.omit$welcome==1, ]) #dropped Education to improve model convergence
summary(m.employed.tot2)

#Question 5: TOT 2.0 Among those that came for a welcome, did uploading a plan
m.employed.plans.tot3 <- lme(employed.pair.delta ~ treat*plans + BYE +  Worker.Profile.Score +
			     Disability,
		     random = ~1 | fo,
		     data = or.dfb.omit2[or.dfb.omit2$welcome==1, ]) #dropped Education to improve model
summary(m.employed.plans.tot3)

#----Only Individuals w/ BYE of 201718 or less [25th percentile]----
#This analysis is to see what the impact is when there is more time for follow-up
m.employed.itt1.LT <- lme(employed~ treat,
			  random = ~1 | fo,
			  data = or.df.LT)
summary(m.employed.itt1.LT)

m.employed.itt2.LT <- lme(employed.pair.delta ~ treat  +
			  Worker.Profile.Score + Education.Level + Disability,
		  random = ~1 | fo,
		  data = or.df.LT.omit)
summary(m.employed.itt2.LT)

m.employed.tot1.LT <- lme(employed.pair.delta ~ treat  +
			  Worker.Profile.Score + Education.Level + Disability,
		  random = ~1 | fo,
		  data = or.df.LT.omit[or.df.LT.omit$welcome==1, ])
summary(m.employed.tot1.LT)

m.plans.emp.tot1.LT <- lme(employed.pair.delta ~ treat*plans  +
			   Worker.Profile.Score + Education.Level + Disability,
		   random = ~1 | fo,
		   data = or.df.LT.omit[or.df.LT.omit$welcome==1, ])
summary(m.plans.emp.tot1.LT)

row1 <- ds.confidence.table(m.exit.itt1b, "UI_Exit_Pair_Centered")
row2 <- ds.confidence.table(m.employed.itt1b, "Employed_Pair_Centered")

knitr::kable(rbind(row1, row2))


```


Then aggregate the data by field office and see if there is a difference as a simple linear model.

```{r Merge_Data2.0}
pairs2$uiexit <- 0
pairs2$emp <- 0
for(i in 1:14)
{
	pairs2[i, "uiexit"] <- mean(or.df.LT[or.df.LT$fo==pairs2[i, "fo.num"], "uiexit"], na.rm = TRUE)
	pairs2[i, "emp"] <- mean(or.df.LT[or.df.LT$fo==pairs2[i, "fo.num"], "employed"], na.rm = TRUE)
}

summary(lm(uiexit ~ treat, data = pairs2))
summary(lm(emp ~ treat, data = pairs2))

```

#Visualization of Outcomes

This is to see the differences between pairs more visually. The second and third plot is only the longer term follow-up.

```{r Plot_Pairs, echo=FALSE}

require(ggplot2)
sbst_palette <- c("#2E9AC4", "#F9834A", "#03CAB9", "#FA5646", "#F2C446", "#999999", "#000000")

pairs2$treat <- as.factor(pairs2$treat)
plot1 <- ggplot(data = pairs2, aes(y = ui.wks, x = pair, shape = treat, color = treat)) + geom_point(size = 6) +
	scale_color_manual(name = "Treatment", values=sbst_palette) +
	scale_shape_manual(name = "Treatment", values= c(17, 16)) +
	ggtitle("Before PEP UI Weeks") +
	ylim(0, 17)
plot1

plot2 <- ggplot(data = pairs2, aes(y = uiexit, x = pair, shape = treat, color = treat)) + geom_point(size = 6) +
	scale_color_manual(name = "Treatment", values=sbst_palette) +
	scale_shape_manual(name = "Treatment", values= c(17, 16)) +
	ggtitle("After PEP UI Exit") +
	ylim(0.00, 0.70)
plot2

plot3 <- ggplot(data = pairs2, aes(y = emp, x = pair, shape = treat, color = treat)) + geom_point(size = 6) +
	scale_color_manual(name = "Treatment", values=sbst_palette) +
	scale_shape_manual(name = "Treatment", values= c(17, 16)) +
	ggtitle("After PEP Employment") +
	ylim(0, 0.10)
plot3

pairs2[, c(1,2,10,12,14,15, 16)]

```

#Survival Analysis
The following analysis looks at the speed at which individuals are exiting UI or entering employment differentially at PEP field offices compared to control offices.

```{r Survival_Analysis}
require(survival)
or.dfb$uiexit.date <- as.Date(or.dfb$`If.Exited,.Date.Claim.Ended`, origin = "1900-01-01")
or.dfb$claim.date <-  as.Date(or.dfb$Initial.Claim.Entered.Date, origin = "1900-01-01")

start.date <- strptime(c("27.03.2016"), format = "%d.%m.%Y")
end.date <- strptime(c("30.09.2016"), format = "%d.%m.%Y")

or.dfb$time.to.exit <- ifelse(or.dfb$uiexit==1, difftime(or.dfb$uiexit.date, or.dfb$claim.date), difftime(end.date, start.date))

or.dfb.omit <- na.omit(or.dfb[c(1, 2, 5, 9, 30, 35, 42, 45, 46:57)])

###################################################################
#----Survival Model------------------------------------------------
m.exit.survival <- coxph(Surv(time.to.exit, uiexit) ~ treat + BYE +
			 WBA + Worker.Profile.Score +
			 cluster(fo), data = or.dfb.omit)
summary(m.exit.survival)

plot(survfit(m.exit.survival), conf.int = TRUE, xlab = "Days", ylab = "Proportion Still on UI")


###############################################################################
#This doesn't work, trying to make survival function showing T vs C Graphically
###############################################################################
#or.treat.df <- with(or.dfb.omit, data.frame(treat =c(0,1),
#                                            BYE=rep(mean(BYE, na.rm = TRUE), 2),
#                                            WBA=rep(mean(WBA, na.rm = TRUE), 2),
#                                            profile = rep(mean(Worker.Profile.Score, na.rm = TRUE), 2)))
#or.treat.df
#
#plot(survfit(m.exit.survival, newdata = or.treat.df), conf.int=TRUE, lty=c(1,2),
#     xlab="Days", ylab = "Proportion on UI")
#
#legend("topright", legend=c("Treat", "Control"),
#       lty=c(1,2), inset=0.02)

```

#Descriptives
Descibe the two groups

```{r Descriptives}
gender <- prop.table(table(or.dfb$Gender.Code, or.dfb$treat), 2) #Proportion of Females and Males by Treatment
age <- c(mean(or.dfb[or.dfb$treat==0, "Age"], na.rm = TRUE), mean(or.dfb[or.dfb$treat==1, "Age"], na.rm = TRUE)) #Age
white <- prop.table(table(or.dfb$Race.White.Flag, or.dfb$treat), 2) #Proportion of White Race by Treatment
aa <- prop.table(table(or.dfb$Race.African.American.Flag, or.dfb$treat), 2) #Proportion of AA by Treatment
disability <- prop.table(table(or.dfb$Disability, or.dfb$treat), 2) #Proportion of Disability by Treatment
veteran <- prop.table(table(or.dfb$Veteran, or.dfb$treat), 2) #Proportion of Veteran by Treatment
reentry <- prop.table(table(or.dfb$Ex.Offender.Flag, or.dfb$treat), 2) #Proportion of Ex Offenders by Treatment
wba <- c(mean(or.dfb[or.dfb$treat==0, "WBA"], na.rm = TRUE), mean(or.dfb[or.dfb$treat==1, "WBA"], na.rm = TRUE)) #weekly benefit amount

desc.table.2 <- rbind(gender, age, white, aa, disability, veteran, reentry, wba)
row.names(desc.table.2) <- c("genderNA", "genderF", "genderM",
			     "age", "nonWhite", "White", "nonAA", "AA",
			     "disabilityNA", "disabilityNR", "disabilityNO", "disabilityYES",
			     "vetNA", "vetOTHER", "vetEligible", "vetVietnam",
			     "reentryNA", "reentryNO", "reentryYES", "WBA")
knitr::kable(desc.table.2)
```

