

#Next Model the Outcome of UI Exit w/o Treatment

First, we deterimined which variables were most predictive of the outcome of interest (exiting UI) in order to remove unnecessary noise to amplify the power of the study.

The variables that were most predictive were **Worker Profile Score**, **Education Level**, and **Disability Status**.

```{r Model_Creation}
#Create a function to predict UI Exit from a given "Variable"
dsMonthlyExitModel <- function(df, variable)
{

	df$ui.exit.rc <- 0
	df[df$Exited.UI.Flag=="Y", "ui.exit.rc"] <- 1
	df$Worker.Profile.Score <- as.numeric(df$Worker.Profile.Score)

	tmp.m1 <- lm(ui.exit.rc ~ df[, variable], data = df)
	print(variable)
	summary(tmp.m1)
	return(tmp.m1)
}

#Create a function to run a model predicting UI Exit from 8 variables
dsMonthlyExitModel2 <- function(df, variable)
{

	df$ui.exit.rc <- 0
	df[df$Exited.UI.Flag=="Y", "ui.exit.rc"] <- 1
	df$Worker.Profile.Score <- as.numeric(df$Worker.Profile.Score)

	tmp.m1 <- lm(ui.exit.rc ~ df[, variable[1]] +
		     df[, variable[2]] +
		     df[, variable[3]] +
		     df[, variable[4]] +
		     df[, variable[5]] +
		     df[, variable[6]] +
		     df[, variable[7]] +
		     df[, variable[8]], data = df)
	print(variable)
	summary(tmp.m1)
	return(tmp.m1)
}

#Create a model to predict UI Exit from 3 variables -- the variables that were predictive in the 8 variable model
dsMonthlyExitModel3 <- function(df, variable)
{

	df$ui.exit.rc <- 0
	df[df$Exited.UI.Flag=="Y", "ui.exit.rc"] <- 1
	df$Worker.Profile.Score <- as.numeric(df$Worker.Profile.Score)

	tmp.m1 <- lm(ui.exit.rc ~ df[, variable[1]] +
		     df[, variable[2]] +
		     df[, variable[3]], data = df)
	print(variable)
	summary(tmp.m1)
	return(tmp.m1)
}


#Use March and April Data first to determine if the 3 variables are significantly predictive
or.df1$ui.exit.rc <- 0
or.df1[or.df1$Exited.UI.Flag=="Y", "ui.exit.rc"] <- 1
or.df1$Worker.Profile.Score <- as.numeric(or.df1$Worker.Profile.Score)
variableList <- c("WBA", "Worker.Profile.Score", "Age", "Education.Level", "Race.White.Flag", "Ethnicity.Hispanic.Latino.Flag", "Ex.Offender.Flag", "Disability")
stats <- c("Estimate", "sigma", "tstat", "pvalue")

modelsAPRIL <- matrix(ncol = 4, nrow = 8, dimnames = list(variableList, stats))
for(i in 1:8)
{
	modelsAPRIL[i, ] <- summary(dsMonthlyExitModel(or.df2, variableList[i]))$coefficient[2, ]
}
knitr::kable(modelsAPRIL) #These are the esimates of the first level of the models for APRIL (because March only had one week of PEP)

#Make a multivariate model with all 8 predictive variables using both March and April
comb.m1 <- (dsMonthlyExitModel2(or.df1, variableList)) #March
comb.m2 <- (dsMonthlyExitModel2(or.df2, variableList)) #April
summary(comb.m1)
summary(comb.m2)

#Use this model going forward, Worker.Profile.Score + Education.Level + Disability
comb.m3 <- dsMonthlyExitModel3(or.df2, variableList[c(2, 4, 8)]) #April
summary(comb.m3) #April

#Run the models for May to see if it holds true -- Yes for Profile and Disability, no for Education
comb.m3.MAY <- dsMonthlyExitModel3(or.df3, variableList[c(2, 4, 8)])
summary(comb.m3.MAY)

```


```{r Pair_Centering}
#----Missing Values so COV adjust in last step now-----
#STEP 1: create covariate adjusted UI exit residuals
#or.dfa <- ds.na.omit(or.df2)
#m1.res <- ds.cov.uiexit(or.dfa)
#summary(m1.res) #check model
#or.dfa$uiexit.resid <- m1.res$residuals #insert residuals into dataframe to be the main outcome

#This is now the first step
#STEP 2: Create pair centered data
ds.create.pairs <- function(df)
{
	df$pair <- as.character(NA)
	df$treat <- 0
	for(i in 1:14)
	{
		df[as.character(df$fo)==pairs2[i, "fo.num"], "pair"] <- as.character(pairs2[i, "pair"])
		df[as.character(df$fo)==pairs2[i, "fo.num"], "treat"] <- pairs2[i, "treat"]
	}

	return(df)
}

or.dfb <- or.df.TOTAL
or.dfb$pair <- as.character(NA)
or.dfb$treat <- 0
for(i in 1:14)
{
	or.dfb[which(as.character(or.dfb$`277.Field.Office`)==pairs2[i, "fo.num"]), "pair"] <- as.character(pairs2[i, "pair"])
	or.dfb[which(as.character(or.dfb$`277.Field.Office`)==pairs2[i, "fo.num"]), "treat"] <- pairs2[i, "treat"]
}

tmp.pair <- c("a", "b", "c", "d", "e", "f", "g")

or.dfb$uiexit.pair.delta <- NA #create a pair variable
or.dfb$employed.pair.delta <- NA

for(i in 1:7)
{
	pair.mean <- mean(or.dfb[or.dfb$pair==tmp.pair[i], "uiexit"], na.rm=TRUE)
	print(pair.mean)
	or.dfb[which(or.dfb$pair==tmp.pair[i]), "uiexit.pair.delta"] <-   or.dfb[which(or.dfb$pair==tmp.pair[i]),"uiexit"] - pair.mean #create pair centered variable

	pair.mean.emp <- mean(or.dfb[or.dfb$pair==tmp.pair[i], "employed"], na.rm = TRUE)
	print(pair.mean.emp)
	or.dfb[which(or.dfb$pair==tmp.pair[i]), "employed.pair.delta"] <- or.dfb[which(or.dfb$pair==tmp.pair[i]), "employed"] - pair.mean.emp
}


```

Save analysis file. Only lose one person here.

```{r}
or.dfb$fo <- as.factor(or.dfb$`277.Field.Office`)
or.dfb <- or.dfb[which(!is.na(or.dfb$PEP.Unique.ID)), ]

save(or.dfb,file="Data/or.dfb.rda")
```
