---
title: "Data Setup"
author: "Daniel Shephard, Jake Bowers, Michael DiDomenico, Paul Testa"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: html_document
---

#Import Data
Create a separate dataframe for each month for the preliminary data analysis from March 2016 through September 2016 (7 months).

```
BYE - benefits end date (201711 is 11th week of 2017)
Inital Claim Entered Date -- when they started
Worker Profile SCore (WPS) - higher more risky of not getting a job
Received Insert - Did you get an insert in the mailing. (Intervention 1)
277 Field Office - Unique IDs for Field Offices in Pairs
Welcome Process Completed Flag - Compliance: they actually had a meeting and made a four week plan and/or they actually made a claim (in the control group) These people actually get their UI payments. Both have to show up in person to get their payment and complete their process.
WBA (weekly benefit amount) -- calculated amount, not actual amount
Exhausted Benefits Flag -- Did you use up all of your benefits
Claimed Previous Week Flag -- Did they make a claim in the week before the data pull
Received UI Previous Week Flag -- did they get paid in the week before the data pull
Previous Four Weeks Average Payment -- Average amount received over the past four weeks. (The average of which ever they have)
Exited UI -- three consecutive weeks of no claims
If Exited Date Claim Ended -- date that those three weeks started. what was first week.
Claim Restarted -- in the ??week?? right before the data pull
Line Flag 45 Date -- If you have caught in a fraudelent claim (so, an indicator that they are reemployed).
Employed Flag -- Is an undercount. Based on reports of wages paid from employers.
(Anyone earning money from an employer not their original employer in a subsequent quarter)
Quarter of Wages --- the first quarter reported, were reported from those Employed (a 5 month delay in the wage file)
WF0.Zip.Code --- perhaps zip code of the home of the person
Email Flag --- do they have an email on file.
Uploaded a Resume -- own or one that was created for them
PEPP Plan Upload Date -- Date uploaded the plan (the staff member uploads the plan after the meeting) Som e staff members uploaded them all up in bulk.
# Pepp Plans -- they should do more plans (mostly people did just one plan)
PEPP Message Sent Date --  they they received the most recent emails, and what the content was, 12 or 14 different emails. Sequential. If they received email 5, then they received email 4.
Pepp message -- message number
# Pepp Messages Sent --- depends on length of time that they are on UI.
```

```{r Data_Import}
library("openxlsx")
filepath <- "Data/PEPP Updated UI Monthly Dataset 030116-123116 NoPassword.xlsx" ## By hand remove the password protection.

#Create dataframes for each month from Mar - December 2016
or.df1 <- read.xlsx(filepath,check.names=TRUE, detectDates=TRUE, sheet = 1) #March
or.df2 <- read.xlsx(filepath,check.names=TRUE, detectDates=TRUE, sheet = 2) #April
or.df3 <- read.xlsx(filepath,check.names=TRUE, detectDates=TRUE, sheet = 3) #May
or.df4 <- read.xlsx(filepath,check.names=TRUE, detectDates=TRUE, sheet = 4) #June
or.df5 <- read.xlsx(filepath,check.names=TRUE, detectDates=TRUE, sheet = 5) #July
or.df6 <- read.xlsx(filepath,check.names=TRUE, detectDates=TRUE, sheet = 6) #August
or.df7 <- read.xlsx(filepath,check.names=TRUE, detectDates=TRUE, sheet = 7) #September
or.df8 <- read.xlsx(filepath,check.names=TRUE, detectDates=TRUE, sheet = 8) #October
or.df9 <- read.xlsx(filepath,check.names=TRUE, detectDates=TRUE, sheet = 9) #November
or.df10 <- read.xlsx(filepath,check.names=TRUE, detectDates=TRUE, sheet = 10) #December
library(lubridate)
or.df10$Initial.Claim.Entered.Date

# Create a merged dataframe from the first 7 months

## Check to make sure all column names are the same
stopifnot(all(names(or.df1)==names(or.df2)))
stopifnot(all(names(or.df1)==names(or.df3)))
stopifnot(all(names(or.df1)==names(or.df4)))
stopifnot(all(names(or.df1)==names(or.df5)))
stopifnot(all(names(or.df1)==names(or.df6)))
stopifnot(all(names(or.df1)==names(or.df7)))
stopifnot(all(names(or.df1)==names(or.df8)))
stopifnot(all(names(or.df1)==names(or.df9)))
stopifnot(all(names(or.df1)==names(or.df10)))



or.df <- rbind(or.df1, or.df2, or.df3, or.df4, or.df5, or.df6, or.df7,or.df8,or.df9,or.df10)
dim(or.df)
stopifnot(nrow(or.df)==length(unique(or.df$PEP.Unique.ID))) #ensure there are no duplicate PEP IDs
```

Check for missing id:

```{r}
## Make sure only 1 person has missing ID data
stopifnot(sum(is.na(or.df$PEP.Unique.ID))==1)
or.df <- or.df[!is.na(or.df$PEP.Unique.ID),]
```

## Duplicates

```{r}
# No duplicate IDs in this dataset
stopifnot(sum(table(or.df$PEP.Unique.ID)>1)==0)
```

## Checking Dates

```{r}
tmp <- table(months(or.df$Initial.Claim.Entered.Date),exclude=c())
stopifnot(nrow(or.df1[!is.na(or.df1$PEP.Unique.ID),])==tmp[["March"]])
stopifnot(nrow(or.df2[!is.na(or.df2$PEP.Unique.ID),])==tmp[["April"]])
stopifnot(nrow(or.df3[!is.na(or.df3$PEP.Unique.ID),])==tmp[["May"]])
stopifnot(nrow(or.df4[!is.na(or.df4$PEP.Unique.ID),])==tmp[["June"]])
stopifnot(nrow(or.df5[!is.na(or.df5$PEP.Unique.ID),])==tmp[["July"]])
stopifnot(nrow(or.df6[!is.na(or.df6$PEP.Unique.ID),])==tmp[["August"]])
stopifnot(nrow(or.df7[!is.na(or.df7$PEP.Unique.ID),])==tmp[["September"]])
stopifnot(nrow(or.df8[!is.na(or.df8$PEP.Unique.ID),])==tmp[["October"]])
stopifnot(nrow(or.df9[!is.na(or.df9$PEP.Unique.ID),])==tmp[["November"]])
stopifnot(nrow(or.df10[!is.na(or.df10$PEP.Unique.ID),])==tmp[["December"]])
```

## Checking Dates

```{r}
range(or.df$Initial.Claim.Entered.Date)

table(or.df$Exited.UI.Flag,useNA="ifany")
# All 23490 who have not exited UI have no date for when claims stopped
sum(is.na(or.df$Date.Claiming.Stopped[or.df$Exited.UI.Flag=="N"]))
# All 23490 who have exited, have a date for when  they exited
sum(!is.na(or.df$Date.Claiming.Stopped[or.df$Exited.UI.Flag=="Y"]))

# Ensure those dates are after the initial claim date
sum(or.df$Date.Claiming.Stopped[or.df$Exited.UI.Flag=="Y"]>or.df$Initial.Claim.Entered.Date[or.df$Exited.UI.Flag=="Y"]) # 22533, 

# 957 have initial claim entered dates after they have claimed to have exited the program, all of these cases are within 30 days of each other

table(or.df$Date.Claiming.Stopped[or.df$Exited.UI.Flag=="Y"]-or.df$Initial.Claim.Entered.Date[or.df$Exited.UI.Flag=="Y"])[1:18]

# 203 have an employed flag
table(or.df$Employed.Flag[or.df$Initial.Claim.Entered.Date>or.df$Date.Claiming.Stopped])

# None have exhausted their benefits
table(or.df$Exhausted.Benefits.Flag[or.df$Initial.Claim.Entered.Date>or.df$Date.Claiming.Stopped])



```


## Examine participates who restart benefits

```{r}
sum(is.na(or.df$Date.Claim.Restarted)) # 23348
sum(!is.na(or.df$Date.Claim.Restarted)) # 17571

# Among thos who Restart Claims, is date after initial claim
sum(or.df$Date.Claim.Restarted>or.df$Initial.Claim.Entered.Date,na.rm=T) # 17571

# None of those re-starting claims had claimed in the previous week
table(or.df$Claimed.Previous.Week.Flag[!is.na(or.df$Date.Claim.Restarted)])
table(or.df$Claimed.Previous.Week.Flag[is.na(or.df$Date.Claim.Restarted)])
# Some have exited UI
table(or.df$Exited.UI.Flag[is.na(or.df$Date.Claim.Restarted)])
table(or.df$Exited.UI.Flag[!is.na(or.df$Date.Claim.Restarted)])
```


## Add row names

```{r}
row.names(or.df) <- or.df$PEP.Unique.ID

```


#### Adding latest wage data

```{r}
or.wages2016Q3 <- read.xlsx("Data/PEPP UI Monthly Dataset Wages Only 030116-093016 No Password.xlsx",check.names=TRUE, detectDates=TRUE)
dim(or.wages2016Q3)

## Make sure that the variable names are the saame
setdiff(names(or.df),names(or.wages2016Q3)) ## names of variables on the main dataset not in the wage update
setdiff(names(or.wages2016Q3),names(or.df))

## Make sure that the IDs are unique
stopifnot(all(table(or.wages2016Q3$PEP.Unique.ID)==1))
row.names(or.wages2016Q3)<-or.wages2016Q3$PEP.Unique.ID

## Make sure that the IDs are the same on both datasets
table(or.wages2016Q3$PEP.Unique.ID %in% row.names(or.df))
table(or.wages2016Q3$PEP.Unique.ID %in% or.df$PEP.Unique.ID)

dim(or.df)

## Update data for people who previously existed in data
existingids <- or.wages2016Q3$PEP.Unique.ID[or.wages2016Q3$PEP.Unique.ID %in% or.df$PEP.Unique.ID]
newids <- or.wages2016Q3$PEP.Unique.ID[!(or.wages2016Q3$PEP.Unique.ID %in% or.df$PEP.Unique.ID)]

stopifnot(length(existingids)+length(newids)==nrow(or.wages2016Q3))

# Check initial claims dates
range(or.wages2016Q3$Initial.Claim.Entered.Date)
stopifnot(all.equal(or.df[existingids,"Initial.Claim.Entered.Date"],or.wages2016Q3[existingids,"Initial.Claim.Entered.Date"]))

#### Fails the next two. 
## We think that the wage file updated all of the other variables on some other date the file was created, not as of Dec 31.
## stopifnot(all.equal(or.df[existingids,"Date.Claiming.Stopped"],or.wages2016Q3[existingids,"Date.Claiming.Stopped"]))
## blah<-or.df[existingids,"Date.Claiming.Stopped"]!=or.wages2016Q3[existingids,"Date.Claiming.Stopped"]
## blah<-or.df[existingids,"Date.Claiming.Stopped"]!=or.wages2016Q3[existingids,"Date.Claiming.Stopped"]
## or.wages2016Q3[blah,c("Date.Claim.Restarted","Date.Claiming.Stopped")][1:10,]
## stopifnot(all.equal(or.df[existingids,"Claim.Restarted.Flag"],or.wages2016Q3[existingids,"Claim.Restarted.Flag"]))
### SO, from the wage file: only keep these variables.
##[19] "Employed.Flag"
##[20] "Quarter.of.Wages"
##[21] "Amount.of.Wages"

summary(or.df[existingids,"Amount.of.Wages"]-or.wages2016Q3[existingids,"Amount.of.Wages"])

## The Wages File is the File of Record??? 

or.df$PEP.Unique.ID
or.df[existingids,c("Employed.Flag","Quarter.of.Wages","Amount.of.Wages")]<-or.wages2016Q3[existingids,c("Employed.Flag","Quarter.of.Wages","Amount.of.Wages")]
# Add new cases
# All newids cases have missig office number, excluded from analysis.
#or.df <- rbind(or.df,or.wages2016Q3[newids,])

stopifnot(all(table(or.df$PEP.Unique.ID)==1))

```

## Recode variables

```{r Recode_Data}
library(Hmisc)
#Set class for variables
or.df$Worker.Profile.Score <- as.numeric(or.df$Worker.Profile.Score)
or.df$Education.Level <- ordered(or.df$Education.Level,
				       levels = c("NONE",  "1ST - 9TH GRADE", "10TH - 12TH GRADE",
						  "HIGH SCHOOL DIPLOMA/GED",
						  "VOCATIONAL CERTIFICATION/DEGREE",
						  "POST SECONDARY COURSEWORK", "ASSOCIATES", "BACHELORS",
						  "MASTERS", "DOCTORATE"))

table(or.df$Employed.Flag, or.df$X277.Field.Office,exclude=c())


# Recode Y / N variables to 1 / 0
or.df$uiexit <- 0 #set variable to 0
or.df[or.df$Exited.UI.Flag=="Y","uiexit"] <- 1

or.df$employed <- 0
or.df[or.df$Employed.Flag=="Y", "employed"] <- 1

or.df$insert <- 0
or.df[or.df$Received.Insert.Flag=="Y", "insert"] <- 1

or.df$welcome <- 0
or.df[which(or.df$Welcome.Process.Completed.Flag=="Y"), "welcome"] <- 1

# Rename dates
or.df$uiexit.date <- or.df$If.Exited..Date.Claim.Ended
or.df$claim.date <-  or.df$Initial.Claim.Entered.Date

```


# Key Analytic Variables

**Units of randomization are field offices** are in `X277.Field.Office`

### Design/Intervention

Pair random assignment of clusters (paired cluster assignment, individual level outcomes)

```{r}
or.df$fonum <- or.df$X277.Field.Office
```

### Outcomes

```{r}
or.df$plans <- as.numeric(or.df$X..of.PEPP.Plans)
or.df$wages <- or.df$Amount.of.Wages

table(or.df$Exhausted.Benefits.Flag)
or.df$exhaust<-ifelse(or.df$Exhausted.Benefits.Flag=="Y",1,0)
table(or.df$Exhausted.Benefits.Flag,or.df$exhaust)
or.df$plans01<-ifelse(or.df$plans>0,1,0)
table(or.df$plans01,useNA="ifany")

# Recode Exit date
table(or.df$uiexit,or.df$exhaust)
or.df$uiexit[or.df$exhaust==1]<-0
table(or.df$uiexit,or.df$exhaust)


or.df$uiexit.date <- as.Date(or.df$Date.Claiming.Stopped, origin = "1900-01-01")
or.df$claim.date <-  as.Date(or.df$Initial.Claim.Entered.Date, origin = "1900-01-01")
range(or.df$claim.date)
start.date <- strptime(c("28.03.2016"), format = "%d.%m.%Y")
end.date <- strptime(c("30.09.2016"), format = "%d.%m.%Y")
#or.df$time.to.exit <- ifelse(or.df$uiexit==1, difftime(or.df$uiexit.date, or.df$claim.date), difftime(end.date, start.date))
# Time to Exit should be difference between 
or.df$time.to.exit <- ifelse(or.df$uiexit==1, difftime(or.df$uiexit.date, or.df$claim.date), difftime(end.date, or.df$claim.date))
range(or.df$time.to.exit)
# 957 have exit dates before intial claims dates before the program began
sum(or.df$time.to.exit<0)
summary(or.df$time.to.exit[or.df$time.to.exit<0])
or.df$month<-month(or.df$Initial.Claim.Entered.Date)
table(or.df$time.to.exit[or.df$time.to.exit<0],or.df$month[or.df$time.to.exit<0])




```

### Covariates

Variables like race and lowest acceptable age, veteran status, etc.. were
entered pre-treatment and in ignorance of treatment status during the filling
out of their online profile.

```{r}
# Race
or.df$white01<-ifelse(or.df$Race.White.Flag=="Y",1,0)
or.df$black01<-ifelse(or.df$Race.African.American.Flag=="Y",1,0)
or.df$asian01<-ifelse(or.df$Race.Asian.Flag=="Y",1,0)
or.df$nhpi01<-ifelse(or.df$Race.Native.Hawaiian.or.Pacific.Islander.Flag=="Y",1,0)
or.df$amerind01<-ifelse(or.df$Race.American.Indian.or.Alaskan.Native.Flag=="Y",1,0)
or.df$unknown01<-ifelse(or.df$Race.Unknown.Flag=="Y",1,0)

# Gender
or.df$female01<-ifelse(or.df$Gender.Code=="F",1,0)
table(or.df$Gender.Code,or.df$female01,useNA="ifany")
# Treat ^ as not female

# Ex felon
or.df$exoffender01<-ifelse(or.df$Ex.Offender.Flag=="Y",1,0)
# Code ^ as nos...
table(or.df$exoffender01,or.df$Ex.Offender.Flag)

# Veteran

table(or.df$Veteran)

# Ethnicity
table(or.df$Ethnicity.Hispanic.Latino.Flag)
or.df$latinx01<-ifelse(or.df$Ethnicity.Hispanic.Latino.Flag=="Y",1,0)

```


## Save the files

```{r}
save(or.df1,or.df2,or.df3,or.df4,or.df5,or.df6,or.df7,or.df8,or.df9,or.df10, file="Data/clean/monthlyordat.rda")
save(or.df, file="Data/clean/or.df.rda")
```

