---
title: "Data Setup"
author: "Daniel Shephard, Jake Bowers, Michael DiDomenico"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: html_document
---

#Import Data
Create a separate dataframe for each month for the preliminary data analysis from March 2016 through September 2016 (7 months).

```
BYE - benefits end date (201711 is 11th week of 2017)
Inital Claim Entered Date -- when they started
Worker Profile SCore (WPS) - higher more risky of not getting a job
Received Insert - Did you get an insert in the mailing. (Intervention 1)
277 Field Office - Unique IDs for Field Offices in Pairs
Welcome Process Completed Flag - Compliance: they actually had a meeting and made a four week plan and/or they actually made a claim (in the control group) These people actually get their UI payments. Both have to show up in person to get their payment and complete their process.
WBA (weekly benefit amount) -- calculated amount, not actual amount
Exhausted Benefits Flag -- Did you use up all of your benefits
Claimed Previous Week Flag -- Did they make a claim in the week before the data pull
Received UI Previous Week Flag -- did they get paid in the week before the data pull
Previous Four Weeks Average Payment -- Average amount received over the past four weeks. (The average of which ever they have)
Exited UI -- three consecutive weeks of no claims
If Exited Date Claim Ended -- date that those three weeks started. what was first week.
Claim Restarted -- in the ??week?? right before the data pull
Line Flag 45 Date -- If you have caught in a fraudelent claim (so, an indicator that they are reemployed).
Employed Flag -- Is an undercount. Based on reports of wages paid from employers.
(Anyone earning money from an employer not their original employer in a subsequent quarter)
Quarter of Wages --- the first quarter reported, were reported from those Employed (a 5 month delay in the wage file)
WF0.Zip.Code --- perhaps zip code of the home of the person
Email Flag --- do they have an email on file.
Uploaded a Resume -- own or one that was created for them
PEPP Plan Upload Date -- Date uploaded the plan (the staff member uploads the plan after the meeting) Som e staff members uploaded them all up in bulk.
# Pepp Plans -- they should do more plans (mostly people did just one plan)
PEPP Message Sent Date --  they they received the most recent emails, and what the content was, 12 or 14 different emails. Sequential. If they received email 5, then they received email 4.
Pepp message -- message number
# Pepp Messages Sent --- depends on length of time that they are on UI.
```


```{r Data_Import}
library("openxlsx")
filepath <- "Data/PEPP UI Monthly Dataset 03282016-09302016 V2 No Password.xlsx" ## By hand remove the password protection.

#Create dataframes for each month from Mar - Sep 2016
or.df1 <- read.xlsx(filepath,check.names=TRUE, detectDates=TRUE, sheet = 1) #March
or.df2 <- read.xlsx(filepath,check.names=TRUE, detectDates=TRUE, sheet = 2) #April
or.df3 <- read.xlsx(filepath,check.names=TRUE, detectDates=TRUE, sheet = 3) #May
or.df4 <- read.xlsx(filepath,check.names=TRUE, detectDates=TRUE, sheet = 4) #June
or.df5 <- read.xlsx(filepath,check.names=TRUE, detectDates=TRUE, sheet = 5) #July
or.df6 <- read.xlsx(filepath,check.names=TRUE, detectDates=TRUE, sheet = 6) #August
or.df7 <- read.xlsx(filepath,check.names=TRUE, detectDates=TRUE, sheet = 7) #September

# Create a merged dataframe from the first 7 months

## Check to make sure all column names are the same
stopifnot(all(names(or.df1)==names(or.df2)))
stopifnot(all(names(or.df1)==names(or.df3)))
stopifnot(all(names(or.df1)==names(or.df4)))
stopifnot(all(names(or.df1)==names(or.df5)))
stopifnot(all(names(or.df1)==names(or.df6)))
stopifnot(all(names(or.df1)==names(or.df7)))

or.df1to7 <- rbind(or.df1, or.df2, or.df3, or.df4, or.df5, or.df6, or.df7)
dim(or.df1to7)
stopifnot(nrow(or.df1to7)==length(unique(or.df1to7$PEP.Unique.ID))) #ensure there are no duplicate PEP IDs
```

Now begin to add other data from other months. For some reason, these files do
not have the same columns as the first months, so we have to rename and strip
off only the columns that matter from **all** of the data in the hopes that we
can make one big file. Not using csv to ease use of dates here. We also had to
replace the variable names for the October file with those from the Nov file
(see `OctVsNovVarNames.csv` for the cross walk of the original names. Of
course, we keep the original files for reference, too.)

```{r}
## or.df9 <- read.csv("Data/PEPP UI Monthly Dataset 110116-113016 No Password.csv")
## Using Excel files here to save work given previous code. In the future should be csv.
or.df9 <- read.xlsx("Data/PEPP UI Monthly Dataset 110116-113016 No Password.xlsx",check.names=TRUE, detectDates=TRUE)
setdiff(names(or.df1),names(or.df9)) ## names of variables on the main dataset not in the 9th set
setdiff(names(or.df9),names(or.df1)) ## names of variables on the 9th dataset not in the main dataset
## Changing some names of variables based on my guesses about equivalences
or.df9$BYE <- or.df9$Claim.BYE
or.df9$PEPP.Plan.Upload.Date <- or.df9$Latest.PEPP.Plan.Upload.Date
or.df9$PEPP.Message <- or.df9$Latest.PEPP.Message.Sent
or.df9$If.Exited..Date.Claim.Ended <- or.df9$Date.Claiming.Stopped
or.df9$PEPP.Message.Sent.Date <- or.df9$Latest.PEPP.Message.Sent.Date
stopifnot(nrow(or.df9)==length(unique(or.df9$PEP.Unique.ID))) #ensure there are no duplicate PEP IDs

## October has the same names as November
or.df8 <- read.xlsx("Data/PEPP UI Monthly Dataset 100116-103116 No Password Nov Names.xlsx",check.names=TRUE, detectDates=TRUE)
setdiff(names(or.df1),names(or.df8)) ## names of variables on the main dataset not in the 8th set
setdiff(names(or.df8),names(or.df1)) ## names of variables on the 8th dataset not in the main dataset
## Changing some names of variables based on my guesses about equivalences
or.df8$BYE <- or.df8$Claim.BYE
or.df8$PEPP.Plan.Upload.Date <- or.df8$Latest.PEPP.Plan.Upload.Date
or.df8$PEPP.Message <- or.df8$Latest.PEPP.Message.Sent
or.df8$If.Exited..Date.Claim.Ended <- or.df8$Date.Claiming.Stopped
or.df8$PEPP.Message.Sent.Date <- or.df8$Latest.PEPP.Message.Sent.Date
stopifnot(nrow(or.df8)==length(unique(or.df8$PEP.Unique.ID))) #ensure there are no duplicate PEP IDs
```

### Combine datasets
```{r}
or.df.TOTAL  <- rbind(or.df1to7,
		      or.df8[,names(or.df1to7)],
		      or.df9[,names(or.df1to7)])
```

### Missing data

Delete the one person with missing id:

```{r}
## Make sure only 1 person has missing ID data
stopifnot(sum(is.na(or.df.TOTAL$PEP.Unique.ID))==1)
or.df.TOTAL <- or.df.TOTAL[!is.na(or.df.TOTAL$PEP.Unique.ID),]
```

### Duplicated IDs

We have some duplicate people even though no duplicates on any given file. Some of these people seem to have entered the system in different months.

```{r}
table(or.df.TOTAL$PEP.Unique.ID)[table(or.df.TOTAL$PEP.Unique.ID)>1]
dupids <- names(table(or.df.TOTAL$PEP.Unique.ID)[table(or.df.TOTAL$PEP.Unique.ID)>1])

dups<-or.df.TOTAL[or.df.TOTAL$PEP.Unique.ID %in% dupids,]
dups <- dups[order(dups$PEP.Unique.ID,dups$Initial.Claim.Entered.Date),]
dups[,c("PEP.Unique.ID","Initial.Claim.Entered.Date",
	"X277.Field.Office","Age","Gender.Code",
	"Exited.UI.Flag")]
```

For now, keeping the first/earlier entry, but we will need actual sleuthing here later.

```{r}

badrows <- sapply(split(dups,dups$PEP.Unique.ID),function(dat){
	       row.names(dat[dat$Initial.Claim.Entered.Date ==
			 max(dat$Initial.Claim.Entered.Date),,drop=FALSE]) })

## Test: given sorting order of the dups data, the later obs should be in even numbered positions
stopifnot(all(badrows==row.names(dups[seq(2,nrow(dups),2),c("PEP.Unique.ID","Initial.Claim.Entered.Date")])))

or.df.TOTAL <- or.df.TOTAL[!(row.names(or.df.TOTAL) %in% badrows),]
stopifnot(all(table(or.df.TOTAL$PEP.Unique.ID)==1))
```

Now make row names of the main dataset equal the ID number:

```{r}
row.names(or.df.TOTAL) <- or.df.TOTAL$PEP.Unique.ID
```

## Check that the date variables in Excel are Dates

```{r}
tmp <- table(months(or.df.TOTAL$Initial.Claim.Entered.Date),exclude=c())
stopifnot(nrow(or.df1[!is.na(or.df1$PEP.Unique.ID),])==tmp[["March"]])
stopifnot(nrow(or.df2[!is.na(or.df2$PEP.Unique.ID),])==tmp[["April"]])
stopifnot(nrow(or.df3[!is.na(or.df3$PEP.Unique.ID),])==tmp[["May"]])
stopifnot(nrow(or.df4[!is.na(or.df4$PEP.Unique.ID),])==tmp[["June"]])
stopifnot(nrow(or.df5[!is.na(or.df5$PEP.Unique.ID),])==tmp[["July"]])
stopifnot(nrow(or.df6[!is.na(or.df6$PEP.Unique.ID),])==tmp[["August"]])
stopifnot(nrow(or.df7[!is.na(or.df7$PEP.Unique.ID),])==tmp[["September"]])
```

#### Adding latest wage data

```{r}
or.wages2016Q3 <- read.xlsx("Data/PEPP UI Monthly Dataset Wages Only 030116-063016 No Password.xlsx",check.names=TRUE, detectDates=TRUE)
setdiff(names(or.df.TOTAL),names(or.wages2016Q3)) ## names of variables on the main dataset not in the wage update
setdiff(names(or.wages2016Q3),names(or.df.TOTAL)) ## names of variables on the wage update not in the main dataset
or.wages2016Q3$PEPP.Plan.Upload.Date <- or.wages2016Q3$Latest.PEPP.Plan.Upload.Date
or.wages2016Q3$PEPP.Message <- or.wages2016Q3$Latest.PEPP.Message.Sent
or.wages2016Q3$If.Exited..Date.Claim.Ended <- or.wages2016Q3$Date.Claiming.Stopped
or.wages2016Q3$PEPP.Message.Sent.Date <- or.wages2016Q3$Latest.PEPP.Message.Sent.Date
or.wages2016Q3<-or.wages2016Q3[,names(or.df.TOTAL)]


## Check for duplicates
stopifnot(all(table(or.wages2016Q3$PEP.Unique.ID)==1))
row.names(or.wages2016Q3)<-or.wages2016Q3$PEP.Unique.ID

## Looks like some of the folks in the updated data are not in the main dataset
table(or.wages2016Q3$PEP.Unique.ID %in% row.names(or.df.TOTAL))
table(or.wages2016Q3$PEP.Unique.ID %in% or.df.TOTAL$PEP.Unique.ID)
## stopifnot(all(or.wages2016Q3$PEP.Unique.ID %in% row.names(or.df.TOTAL)))

## Update data for people who previously existed in data
existingids <- or.wages2016Q3$PEP.Unique.ID[or.wages2016Q3$PEP.Unique.ID %in% or.df.TOTAL$PEP.Unique.ID]
newids <- or.wages2016Q3$PEP.Unique.ID[!(or.wages2016Q3$PEP.Unique.ID %in% or.df.TOTAL$PEP.Unique.ID)]

stopifnot(length(existingids)+length(newids)==nrow(or.wages2016Q3))

or.df.TOTAL[existingids,]<-or.wages2016Q3[existingids,]

## Add rows to the dataset for people who previously had not bee in the data.

or.df.TOTAL <- rbind(or.df.TOTAL,or.wages2016Q3[newids,])

stopifnot(all(table(or.df.TOTAL$PEP.Unique.ID)==1))

```


## Recode variables

```{r Recode_Data}
library(Hmisc)
#Set class for variables
or.df.TOTAL$Worker.Profile.Score <- as.numeric(or.df.TOTAL$Worker.Profile.Score)
or.df.TOTAL$Education.Level <- ordered(or.df.TOTAL$Education.Level,
				       levels = c("NONE",  "1ST - 9TH GRADE", "10TH - 12TH GRADE",
						  "HIGH SCHOOL DIPLOMA/GED",
						  "VOCATIONAL CERTIFICATION/DEGREE",
						  "POST SECONDARY COURSEWORK", "ASSOCIATES", "BACHELORS",
						  "MASTERS", "DOCTORATE"))

table(or.df.TOTAL$Employed.Flag, or.df.TOTAL$X277.Field.Office,exclude=c())

# Recode Y / N variables to 1 / 0
or.df.TOTAL$uiexit <- 0 #set variable to 0
or.df.TOTAL[or.df.TOTAL$Exited.UI.Flag=="Y","uiexit"] <- 1

or.df.TOTAL$employed <- 0
or.df.TOTAL[or.df.TOTAL$Employed.Flag=="Y", "employed"] <- 1

or.df.TOTAL$insert <- 0
or.df.TOTAL[or.df.TOTAL$Received.Insert.Flag=="Y", "insert"] <- 1

or.df.TOTAL$welcome <- 0
or.df.TOTAL[which(or.df.TOTAL$Welcome.Process.Completed.Flag=="Y"), "welcome"] <- 1

# Rename dates
or.df.TOTAL$uiexit.date <- or.df.TOTAL$If.Exited..Date.Claim.Ended
or.df.TOTAL$claim.date <-  or.df.TOTAL$Initial.Claim.Entered.Date
```


# Key Analytic Variables

**Units of randomization are field offices** are in `X277.Field.Office`

### Design/Intervention

Pair random assignment of clusters (paired cluster assignment, individual level outcomes)

```{r}
or.df.TOTAL$fonum <- or.df.TOTAL$X277.Field.Office
```

### Outcomes

```{r}
or.df.TOTAL$plans <- as.numeric(or.df.TOTAL$X..of.PEPP.Plans)
or.df.TOTAL$wages <- or.df.TOTAL$Amount.of.Wages
or.df.TOTAL$uiexit.date <- as.Date(or.df.TOTAL$If.Exited..Date.Claim.Ended, origin = "1900-01-01")
or.df.TOTAL$claim.date <-  as.Date(or.df.TOTAL$Initial.Claim.Entered.Date, origin = "1900-01-01")
start.date <- strptime(c("27.03.2016"), format = "%d.%m.%Y")
end.date <- strptime(c("30.09.2016"), format = "%d.%m.%Y")
or.df.TOTAL$time.to.exit <- ifelse(or.df.TOTAL$uiexit==1, difftime(or.df.TOTAL$uiexit.date, or.df.TOTAL$claim.date), difftime(end.date, start.date))
```

### Covariates

Variables like race and lowest acceptable age, veteran status, etc.. were
entered pre-treatment and in ignorance of treatment status during the filling
out of their online profile.

## Save the files

```{r}
save(or.df1,or.df2,or.df3,or.df4,or.df5,or.df6,or.df7,or.df8,or.df9,
     file="Data/monthlyordat.rda")
save(or.df.TOTAL, file="Data/or.df.TOTAL.rda")
```

